---
title: "Population-based cohorts in Africa: One Health scoping review"
author: "Lydiah Kisoo"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
    toc_depth: '3'
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

# Overview

This R Markdown document contains the reproducible workflow used for the
scoping review of population-based cohorts in Africa with relevance to
One Health research.

The workflow includes:

- bibliographic-record cleaning and deduplication;
- cleaning and harmonisation of extracted cohort data;
- construction of human, animal and environmental domains;
- descriptive cohort characteristics;
- infectious-disease outcome summaries;
- laboratory assay and biospecimen summaries;
- data-integration scoring;
- multivariable Poisson regression;
- temporal analyses;
- geographic analyses; and
- recruitment-strategy summaries.

The analytical unit is the **cohort**, rather than the
harmonised cohort-name string. Some publications or networks report more
than one cohort/site together. Therefore, a unique `cohort_id` is used for
cohort-level counts and denominators.

# 1. Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 300
)

options(stringsAsFactors = FALSE)

library(readxl)
library(readr)
library(revtools)
library(stringdist)
library(dplyr)
library(stringr)
library(tidyr)
library(purrr)
library(tibble)
library(forcats)
library(ggplot2)
library(scales)
library(flextable)
library(officer)
library(sandwich)
library(lmtest)
library(countrycode)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggrepel)
library(ggspatial)
```

# 2. Bibliographic search cleaning and deduplication

Search results from the databases were combined before deduplication.
Titles were first converted to UTF-8 and likely duplicate records were
identified using fuzzy title matching.

```{r search-deduplication}
# All input files are expected to be in the same folder as this R Markdown file.

required_search_file <- "combined search.csv"

if (!file.exists(required_search_file)) {
  stop(
    "Cannot find 'combined search.csv'. ",
    "Place the file in the same Systematic Review folder as this R Markdown file."
  )
}

# Import combined bibliographic search results
combineddata <- read.csv(
  "combined search.csv",
  stringsAsFactors = FALSE,
  check.names = FALSE
)

cat("Records imported:", nrow(combineddata), "\n")
```

## 2.1 Clean text encoding

```{r clean-search-text}
cleaneddataset <- combineddata %>%
  mutate(
    Title = iconv(
      Title,
      from = "",
      to = "UTF-8",
      sub = "byte"
    )
  )

write.csv(
  cleaneddataset,
  "search_records_encoding_cleaned.csv",
  row.names = FALSE
)
```

## 2.2 Identify likely duplicates

Potential duplicate titles were identified using optimal string alignment
with a string-distance threshold of five.

```{r identify-duplicates}
duplicate_group <- revtools::find_duplicates(
  data = cleaneddataset,
  match_variable = "Title",
  match_function = "stringdist",
  method = "osa",
  threshold = 5,
  to_lower = TRUE,
  remove_punctuation = TRUE
)

search_with_duplicate_group <- cleaneddataset %>%
  mutate(
    duplicate_group = duplicate_group
  )

potential_duplicates <- search_with_duplicate_group %>%
  group_by(duplicate_group) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(duplicate_group)

write.csv(
  potential_duplicates,
  "potential_duplicate_records.csv",
  row.names = FALSE
)

cat(
  "Records in potential duplicate groups:",
  nrow(potential_duplicates),
  "\n"
)
```

The potential duplicate groups were manually reviewed before the final
deduplicated database was used for screening.

```{r final-deduplication}
deduplicated_search <- search_with_duplicate_group %>%
  group_by(duplicate_group) %>%
  slice(1L) %>%
  ungroup()

write.csv(
  deduplicated_search,
  "search_records_deduplicated.csv",
  row.names = FALSE
)

cat(
  "Records before deduplication:",
  nrow(cleaneddataset),
  "\n"
)

cat(
  "Records after deduplication:",
  nrow(deduplicated_search),
  "\n"
)
```

# 3. Load extracted cohort data

The following datasets contain information extracted after title/abstract
and full-text screening.

```{r load-extracted-data}
required_extraction_files <- c(
  "cohort_info2.xlsx",
  "oh_outcomes2.xlsx",
  "oh_exposures2.xlsx",
  "outcome_details2.xlsx"
)

missing_extraction_files <- required_extraction_files[
  !file.exists(required_extraction_files)
]

if (length(missing_extraction_files) > 0) {
  stop(
    "The following required files are missing from the current folder: ",
    paste(missing_extraction_files, collapse = ", ")
  )
}

cohort_info <- read_excel(
  "cohort_info2.xlsx"
)

oh_outcomes <- read_excel(
  "oh_outcomes2.xlsx"
)

oh_exposures <- read_excel(
  "oh_exposures2.xlsx"
)

outcome_details <- read_excel(
  "outcome_details2.xlsx"
)

cat("Cohort information rows:", nrow(cohort_info), "\n")
cat("Outcome rows:", nrow(oh_outcomes), "\n")
cat("Exposure rows:", nrow(oh_exposures), "\n")
cat("Outcome-detail rows:", nrow(outcome_details), "\n")
```

# 4. Data-cleaning functions

```{r helper-functions}
clean_text <- function(x) {

  x <- as.character(x)

  x <- str_replace_all(
    x,
    "\u2019|\u2018",
    "'"
  )

  x <- str_replace_all(
    x,
    "\u2013|\u2014",
    "-"
  )

  x <- str_replace_all(
    x,
    "\u03b2",
    "beta"
  )

  str_squish(x)
}


to_yesno_strict <- function(x) {

  if (is.numeric(x)) {

    return(
      ifelse(
        !is.na(x) & x == 1,
        "Yes",
        "No"
      )
    )
  }

  x <- tolower(
    trimws(
      as.character(x)
    )
  )

  ifelse(
    x %in% c(
      "yes",
      "y",
      "1",
      "true",
      "present"
    ),
    "Yes",
    "No"
  )
}


# Used for exposure variables:
# any substantive non-empty response indicates assessment/presence.

to_yesno_exposure <- function(x) {

  x_chr <- trimws(
    as.character(x)
  )

  x_low <- tolower(x_chr)

  x_low <- str_replace_all(
    x_low,
    "\\s+",
    " "
  )

  no_patterns <- c(
    "",
    "no",
    "n",
    "0",
    "false",
    "absent",
    "none",
    "not assessed",
    "none assessed",
    "not reported",
    "na",
    "n/a"
  )

  ifelse(
    !is.na(x_chr) &
      !(x_low %in% no_patterns),
    "Yes",
    "No"
  )
}


collapse_unique_text <- function(
    x,
    sep = "; "
) {

  x <- clean_text(x)

  x <- x[
    !is.na(x) &
      x != "" &
      tolower(x) != "none" &
      tolower(x) != "not specified"
  ]

  x <- unique(x)

  if (length(x) == 0) {
    return(NA_character_)
  }

  paste(
    x,
    collapse = sep
  )
}


flag_detect <- function(
    x,
    pattern
) {

  x <- clean_text(x)

  ifelse(
    any(
      str_detect(
        str_to_lower(x),
        pattern
      ),
      na.rm = TRUE
    ),
    "Yes",
    "No"
  )
}


yes_count <- function(
    df,
    cols
) {

  if (length(cols) == 0) {

    rep(
      0L,
      nrow(df)
    )

  } else {

    rowSums(
      df[, cols, drop = FALSE] == "Yes",
      na.rm = TRUE
    )
  }
}


force_yesno <- function(x) {

  ifelse(
    !is.na(x) &
      as.character(x) == "Yes",
    "Yes",
    "No"
  )
}
```

# 5. Harmonise cohort names

Some cohort names differed across extraction sheets. A manual crosswalk was
therefore used to harmonise names before merging datasets.

Importantly, harmonisation is used for **linking records across extraction
sheets**. The harmonised name is not subsequently assumed to be a unique
cohort identifier because some networks contain multiple distinct cohort
sites.

```{r cohort-name-crosswalk}
name_crosswalk <- tibble::tribble(

  ~raw_name,
  ~harmonized_name,

  "Manyatta HDSS",
  "Manyatta Karemo HDSS",

  "Karemo HDSS",
  "Manyatta Karemo HDSS",

  "Manyatta/Karemo HDSS",
  "Manyatta Karemo HDSS",

  "Kilifi HDSS2",
  "Kilifi HDSS",

  "Kisumu Ahero/KEMRI CDC",
  "Kisumu KEMRI CDC",

  "Dinamo HDSS/Dikgale",
  "Dinamo HDSS",

  "MLSFH/gen-malawi",
  "MLSFH",

  "Djicoroni Para&Banconi, Bamako_CHAMPS_Network",
  "Harar_Djicoroni_CHAMPS",

  "Harar_CHAMPS_Network",
  "Harar_Djicoroni_CHAMPS",

  paste0(
    "GRT-INSPIRED HDSS (Gauteng Research Triangle Initiative ",
    "for the Study of Population, Infrastructure and Regional ",
    "Economic Development)"
  ),
  "GRT-INSPIRED HDSS",

  "GRT-INSPIRED HDSS (Gauteng)",
  "GRT-INSPIRED HDSS",

  paste0(
    "C-SHARP HDSS (Cape Town Surveillance through ",
    "Healthcare Action Research Project)"
  ),
  "C-SHARP HDSS",

  "C-SHARP HDSS (Western Cape)",
  "C-SHARP HDSS",

  "BAMMISHO",
  "BAMMISHO HDSS",

  paste0(
    "Northwestern Tigray Rabies Exposure Cohort ",
    "(Suhul Hospital Catchment)"
  ),
  "Northwestern Tigray Rabies Exposure Cohort",

  paste0(
    "PATHOME Study (Pathogen Transmission and Health ",
    "Outcome Models of Enteric Disease)"
  ),
  "PATHOME Study",

  "Dielmo & Ndiop Malaria Cohort",
  "Dielmo and Ndiop Malaria Cohorts (Senegal)"
)


harmonize_cohort_name <- function(
    x,
    crosswalk = name_crosswalk
) {

  x_clean <- clean_text(x)

  matched <- crosswalk$harmonized_name[
    match(
      x_clean,
      crosswalk$raw_name
    )
  ]

  ifelse(
    is.na(matched),
    x_clean,
    matched
  )
}
```

Apply the harmonisation consistently across all four extraction datasets.

```{r apply-name-harmonisation}
cohort_info <- cohort_info %>%
  mutate(
    cohort_name_raw = clean_text(cohort_name),
    cohort_name = harmonize_cohort_name(
      cohort_name_raw
    )
  )


oh_outcomes <- oh_outcomes %>%
  mutate(
    cohort_name_raw = clean_text(cohort_name),
    cohort_name = harmonize_cohort_name(
      cohort_name_raw
    )
  )


oh_exposures <- oh_exposures %>%
  mutate(
    cohort_name_raw = clean_text(cohort_name),
    cohort_name = harmonize_cohort_name(
      cohort_name_raw
    )
  )


outcome_details <- outcome_details %>%
  mutate(
    cohort_name_raw = clean_text(cohort_name),
    cohort_name = harmonize_cohort_name(
      cohort_name_raw
    )
  )
```

# 6. Clean outcome variables

Two Brucellosis columns were created during import because the original
extraction sheet contained duplicate column names. These are combined into
one indicator.

```{r brucellosis-cleaning}
if (
  all(
    c(
      "Brucellosis...30",
      "Brucellosis...36"
    ) %in% names(oh_outcomes)
  )
) {

  oh_outcomes <- oh_outcomes %>%

    mutate(

      Brucellosis = ifelse(

        to_yesno_strict(
          `Brucellosis...30`
        ) == "Yes" |

          to_yesno_strict(
            `Brucellosis...36`
          ) == "Yes",

        "Yes",
        "No"
      )

    ) %>%

    select(
      -`Brucellosis...30`,
      -`Brucellosis...36`
    )
}
```

Collapse repeated outcome records to the harmonised cohort level.

```{r collapse-outcomes}
outcome_cols_raw <- setdiff(
  names(oh_outcomes),
  c(
    "cohort_name",
    "cohort_name_raw"
  )
)


oh_outcomes_cohort <- oh_outcomes %>%

  mutate(
    across(
      all_of(outcome_cols_raw),
      to_yesno_strict
    )
  ) %>%

  group_by(cohort_name) %>%

  summarise(

    across(
      all_of(outcome_cols_raw),
      ~ ifelse(
        any(.x == "Yes"),
        "Yes",
        "No"
      )
    ),

    .groups = "drop"
  )


outcome_list_by_cohort <- oh_outcomes_cohort %>%

  rowwise() %>%

  mutate(

    outcomes_list = paste(

      names(
        c_across(
          all_of(outcome_cols_raw)
        )
      )[
        c_across(
          all_of(outcome_cols_raw)
        ) == "Yes"
      ],

      collapse = "; "
    ),

    outcomes_list = ifelse(
      outcomes_list == "",
      NA_character_,
      outcomes_list
    )
  ) %>%

  ungroup() %>%

  select(
    cohort_name,
    outcomes_list
  )
```

# 7. Define One Health exposure groups

```{r exposure-variable-groups}
animal_vars <- c(
  "Wildlife/wildbird Proximity or contact",
  "Livestock ownership (general)",
  "Livestock ownership (Pig)",
  "Poultry/birds Ownership",
  "Livestock ownership (Goat)",
  "Livestock ownership (Cattle)",
  "Livestock ownership (Sheep)",
  "Small animals",
  "Bats contact",
  "Bushmeat consumption",
  "Cat contact/ownership",
  "Dog contact/bites",
  "Animal Sampling",
  "Animal Health Practices(Surveys - Biosecurity, vaccination)",
  "Antibiotic use in animals",
  "Dog Vaccination"
)


ecosystem_vars <- c(
  "WASH (Hand washing, saniation, access to clean water)",
  "Natural water body contact",
  "Water waste management",
  "Solid waste management",
  "WASH(Toilet type)",
  "Food secutity (nutrition)",
  "Air pollution(Particulate Matter (PM2.5 or other)",
  "Air pollution(Energy Source Cooking)",
  "Air Pollution(Dust exposure, proximity to Road etc)",
  "Seasonality",
  "Climate Variables(Temp.)",
  "Climate Variables(Humidity)",
  "Climate Variables(Dryness)",
  "Climate Variables(Rainfall)",
  "Greenness",
  "Noise Pollution",
  "Chemical exposure",
  "Flooding context",
  "Soil contact / geophagy",
  "Other_Environmental sampling"
)


vector_vars <- c(
  "Use of bed nets",
  "Entomological survey - Mosquitoes(densities etc)",
  "Entomological survey - Ticks",
  "Vector exposure(mosquitoes, snails, ticks)"
)


socioeconomic_vars <- c(
  "Socio-economic Indicators"
)


other_contextual_vars <- c(
  "Geolocation (GPS)",
  "Residency(Rural/Urban)"
)
```

# 8. Derive occupation-related exposure indicators

```{r occupation-exposures}
occupation_animal_pattern <- paste(
  c(
    "pastoral",
    "poultry",
    "livestock",
    "hunting",
    "hunting and farming",
    "farming and pastoralism"
  ),
  collapse = "|"
)


occupation_ecosystem_pattern <- paste(
  c(
    "farming",
    "agriculture",
    "fishermen",
    "fishing",
    "mining",
    "pastoral",
    "hunting",
    "poultry"
  ),
  collapse = "|"
)


oh_exposures2 <- oh_exposures %>%

  mutate(

    occupation_value = str_to_lower(
      clean_text(
        `Occupation_ OH related (Specify)`
      )
    ),

    occupation_animal_related = ifelse(
      str_detect(
        occupation_value,
        occupation_animal_pattern
      ),
      "Yes",
      "No"
    ),

    occupation_ecosystem_related = ifelse(
      str_detect(
        occupation_value,
        occupation_ecosystem_pattern
      ),
      "Yes",
      "No"
    )
  )
```

Only variables actually present in the extraction sheet are retained.

```{r available-exposure-columns}
animal_cols <- unique(
  c(
    intersect(
      animal_vars,
      names(oh_exposures2)
    ),
    "occupation_animal_related"
  )
)


ecosystem_cols <- unique(
  c(
    intersect(
      ecosystem_vars,
      names(oh_exposures2)
    ),
    "occupation_ecosystem_related"
  )
)


vector_cols <- intersect(
  vector_vars,
  names(oh_exposures2)
)


socioeconomic_cols <- intersect(
  socioeconomic_vars,
  names(oh_exposures2)
)


other_contextual_cols <- intersect(
  other_contextual_vars,
  names(oh_exposures2)
)


all_exposure_cols <- unique(
  c(
    animal_cols,
    ecosystem_cols,
    vector_cols,
    socioeconomic_cols,
    other_contextual_cols
  )
)
```

Collapse exposure variables to cohort level.

```{r collapse-exposures}
oh_exposures_cohort <- oh_exposures2 %>%

  mutate(
    across(
      all_of(all_exposure_cols),
      to_yesno_exposure
    )
  ) %>%

  group_by(cohort_name) %>%

  summarise(

    across(
      all_of(all_exposure_cols),
      ~ ifelse(
        any(.x == "Yes"),
        "Yes",
        "No"
      )
    ),

    .groups = "drop"
  )
```

# 9. Initial One Health domain indicators

```{r initial-domain-indicators}
cohort_domain_flags <- oh_exposures_cohort %>%

  mutate(

    human_outcome = "Yes",

    animal_yes_n = yes_count(
      .,
      animal_cols
    ),

    ecosystem_yes_n = yes_count(
      .,
      ecosystem_cols
    ),

    vector_yes_n = yes_count(
      .,
      vector_cols
    ),

    socioeconomic_yes_n = yes_count(
      .,
      socioeconomic_cols
    ),

    other_contextual_yes_n = yes_count(
      .,
      other_contextual_cols
    ),

    animal_exposure_domain = ifelse(
      animal_yes_n > 0,
      "Yes",
      "No"
    ),

    ecosystem_exposure_domain = ifelse(
      ecosystem_yes_n > 0,
      "Yes",
      "No"
    ),

    vector_exposure_domain = ifelse(
      vector_yes_n > 0,
      "Yes",
      "No"
    ),

    socioeconomic_exposure_domain = ifelse(
      socioeconomic_yes_n > 0,
      "Yes",
      "No"
    ),

    other_contextual_exposure_domain = ifelse(
      other_contextual_yes_n > 0,
      "Yes",
      "No"
    )
  ) %>%

  select(
    cohort_name,
    human_outcome,
    animal_exposure_domain,
    ecosystem_exposure_domain,
    vector_exposure_domain,
    socioeconomic_exposure_domain,
    other_contextual_exposure_domain
  )
```