---
title: "Population-based cohorts in Africa: One Health scoping review"
author: "Lydiah Kisoo"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
    toc_depth: 3
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

# Overview

This R Markdown document contains the reproducible workflow used for the
scoping review of population-based cohorts in Africa with relevance to
One Health research.

The analytical unit is the cohort/site record as some publications described more than one cohort conducted in different sites in the same manner. Harmonised cohort names are
used to link extraction sheets, while `cohort_id` is used for cohort-level
counts and denominators because some publications or networks report more
than one cohort/site together.

# 1. Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 300
)

options(stringsAsFactors = FALSE)
```

```{r packages}
library(readxl)
library(readr)
library(revtools)
library(stringdist)
library(dplyr)
library(stringr)
library(tidyr)
library(purrr)
library(tibble)
library(forcats)
library(ggplot2)
library(scales)
library(flextable)
library(officer)
library(sandwich)
library(lmtest)
library(countrycode)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggrepel)
library(ggspatial)
```

# 2. Bibliographic search cleaning and deduplication

All files are expected to be stored in the same folder as this R Markdown
file.

```{r search-deduplication}
if (!file.exists("combined search.csv")) {
  stop("Cannot find 'combined search.csv' in the current folder.")
}

combineddata <- read.csv(
  "combined search.csv",
  stringsAsFactors = FALSE,
  check.names = FALSE
)

cat("Records imported:", nrow(combineddata), "\n")

cleaneddataset <- combineddata %>%
  mutate(
    Title = iconv(
      Title,
      from = "",
      to = "UTF-8",
      sub = "byte"
    )
  )

write.csv(
  cleaneddataset,
  "search_records_encoding_cleaned.csv",
  row.names = FALSE
)

duplicate_group <- revtools::find_duplicates(
  data = cleaneddataset,
  match_variable = "Title",
  match_function = "stringdist",
  method = "osa",
  threshold = 5,
  to_lower = TRUE,
  remove_punctuation = TRUE
)

search_with_duplicate_group <- cleaneddataset %>%
  mutate(
    duplicate_group = duplicate_group
  )

potential_duplicates <- search_with_duplicate_group %>%
  group_by(duplicate_group) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(duplicate_group)

write.csv(
  potential_duplicates,
  "potential_duplicate_records.csv",
  row.names = FALSE
)

# Final duplicate groups were manually reviewed during the review process.
deduplicated_search <- search_with_duplicate_group %>%
  group_by(duplicate_group) %>%
  slice(1L) %>%
  ungroup()

write.csv(
  deduplicated_search,
  "search_records_deduplicated.csv",
  row.names = FALSE
)

cat(
  "Records before deduplication:",
  nrow(cleaneddataset),
  "\n"
)

cat(
  "Records after deduplication:",
  nrow(deduplicated_search),
  "\n"
)
```

# 3. Load extracted cohort data

```{r load-extracted-data}
required_extraction_files <- c(
  "cohort_info2.xlsx",
  "oh_outcomes2.xlsx",
  "oh_exposures2.xlsx",
  "outcome_details2.xlsx"
)

missing_extraction_files <- required_extraction_files[
  !file.exists(required_extraction_files)
]

if (length(missing_extraction_files) > 0) {
  stop(
    "Missing required file(s): ",
    paste(
      missing_extraction_files,
      collapse = ", "
    )
  )
}

cohort_info <- read_excel(
  "cohort_info2.xlsx"
)

oh_outcomes <- read_excel(
  "oh_outcomes2.xlsx"
)

oh_exposures <- read_excel(
  "oh_exposures2.xlsx"
)

outcome_details <- read_excel(
  "outcome_details2.xlsx"
)

cat(
  "Cohort information rows:",
  nrow(cohort_info),
  "\n"
)

cat(
  "Outcome rows:",
  nrow(oh_outcomes),
  "\n"
)

cat(
  "Exposure rows:",
  nrow(oh_exposures),
  "\n"
)

cat(
  "Outcome-detail rows:",
  nrow(outcome_details),
  "\n"
)
```

# 4. Helper functions

```{r helper-functions}
clean_text <- function(x) {

  x <- as.character(x)

  x <- str_replace_all(
    x,
    "\u2019|\u2018",
    "'"
  )

  x <- str_replace_all(
    x,
    "\u201c|\u201d",
    '"'
  )

  x <- str_replace_all(
    x,
    "\u2013|\u2014",
    "-"
  )

  x <- str_replace_all(
    x,
    "\u03b2",
    "beta"
  )

  str_squish(x)
}


to_yesno_strict <- function(x) {

  if (is.numeric(x)) {

    return(
      ifelse(
        !is.na(x) & x == 1,
        "Yes",
        "No"
      )
    )
  }

  x <- tolower(
    trimws(
      as.character(x)
    )
  )

  ifelse(
    x %in% c(
      "yes",
      "y",
      "1",
      "true",
      "present"
    ),
    "Yes",
    "No"
  )
}


to_yesno_exposure <- function(x) {

  x_chr <- trimws(
    as.character(x)
  )

  x_low <- str_replace_all(
    tolower(x_chr),
    "\\s+",
    " "
  )

  no_patterns <- c(
    "",
    "no",
    "n",
    "0",
    "false",
    "absent",
    "none",
    "not assessed",
    "none assessed",
    "not reported",
    "na",
    "n/a"
  )

  ifelse(
    !is.na(x_chr) &
      !(x_low %in% no_patterns),
    "Yes",
    "No"
  )
}


collapse_unique_text <- function(
    x,
    sep = "; "
) {

  x <- clean_text(x)

  x <- x[
    !is.na(x) &
      x != "" &
      tolower(x) != "none" &
      tolower(x) != "not specified"
  ]

  x <- unique(x)

  if (length(x) == 0) {
    return(NA_character_)
  }

  paste(
    x,
    collapse = sep
  )
}


flag_detect <- function(
    x,
    pattern
) {

  x <- clean_text(x)

  ifelse(
    any(
      str_detect(
        str_to_lower(x),
        pattern
      ),
      na.rm = TRUE
    ),
    "Yes",
    "No"
  )
}


yes_count <- function(
    df,
    cols
) {

  cols <- intersect(
    cols,
    names(df)
  )

  if (length(cols) == 0) {

    rep(
      0L,
      nrow(df)
    )

  } else {

    rowSums(
      df[, cols, drop = FALSE] == "Yes",
      na.rm = TRUE
    )
  }
}


force_yesno <- function(x) {

  ifelse(
    !is.na(x) &
      as.character(x) == "Yes",
    "Yes",
    "No"
  )
}


is_yes <- function(x) {

  x2 <- tolower(
    clean_text(x)
  )

  case_when(

    is.na(x2) ~ FALSE,

    x2 %in% c(
      "1",
      "yes",
      "y",
      "true",
      "present",
      "reported"
    ) ~ TRUE,

    x2 %in% c(
      "0",
      "no",
      "n",
      "false",
      "absent",
      "not reported",
      "",
      "na",
      "n/a"
    ) ~ FALSE,

    suppressWarnings(
      !is.na(as.numeric(x2)) &
        as.numeric(x2) > 0
    ) ~ TRUE,

    TRUE ~ FALSE
  )
}


clean_country_type <- function(x) {

  x <- clean_text(x)

  case_when(

    is.na(x) | x == "" ~
      NA_character_,

    str_detect(
      x,
      regex(
        "^multinational",
        ignore_case = TRUE
      )
    ) ~ "Multi-country",

    str_detect(
      x,
      regex(
        paste0(
          "multiple countries|multi-country|multi country|",
          "[0-9]+\\s+sites|[0-9]+\\s+countries|",
          "african sites include"
        ),
        ignore_case = TRUE
      )
    ) ~ "Multi-country",

    str_count(x, ",") >= 1 ~
      "Multi-country",

    str_count(x, ";") >= 1 ~
      "Multi-country",

    str_detect(
      x,
      regex(
        "\\band\\b",
        ignore_case = TRUE
      )
    ) ~ "Multi-country",

    TRUE ~ "Single country"
  )
}


classify_region <- function(
    country_clean
) {

  case_when(

    is.na(country_clean) |
      country_clean == "" ~
      NA_character_,

    clean_country_type(
      country_clean
    ) == "Multi-country" ~
      "Multinational",

    str_detect(
      country_clean,
      regex(
        paste0(
          "kenya|uganda|tanzania|ethiopia|sudan|",
          "south sudan|rwanda|burundi|eritrea|",
          "somalia|djibouti|seychelles"
        ),
        ignore_case = TRUE
      )
    ) ~ "East Africa",

    str_detect(
      country_clean,
      regex(
        paste0(
          "ghana|nigeria|gambia|the gambia|senegal|",
          "mali|burkina faso|niger|benin|",
          "côte d'ivoire|cote d'ivoire|ivory coast|",
          "guinea-bissau|guinea|liberia|sierra leone|togo"
        ),
        ignore_case = TRUE
      )
    ) ~ "West Africa",

    str_detect(
      country_clean,
      regex(
        paste0(
          "south africa|zimbabwe|zambia|mozambique|",
          "malawi|namibia|botswana|angola|madagascar|",
          "mauritius|lesotho|eswatini"
        ),
        ignore_case = TRUE
      )
    ) ~ "Southern Africa",

    str_detect(
      country_clean,
      regex(
        paste0(
          "democratic republic of congo|",
          "democratic republic of the congo|",
          "dr congo|drc|republic of congo|",
          "central african republic|cameroon|",
          "chad|gabon|equatorial guinea"
        ),
        ignore_case = TRUE
      )
    ) ~ "Central Africa",

    str_detect(
      country_clean,
      regex(
        "egypt|algeria|morocco|tunisia|libya",
        ignore_case = TRUE
      )
    ) ~ "North Africa",

    TRUE ~ NA_character_
  )
}


extract_sample_size_numeric <- function(x) {

  x <- clean_text(x)

  if (
    is.na(x) ||
      x == ""
  ) {
    return(NA_real_)
  }

  x2 <- x %>%

    str_replace_all(
      "(?<=\\d)\\.(?=\\d{3}(\\D|$))",
      ""
    ) %>%

    str_replace_all(
      ",",
      ""
    )

  m1 <- str_match(
    x2,
    regex(
      "humans?\\s*:?\\s*([0-9]+)",
      ignore_case = TRUE
    )
  )

  if (!is.na(m1[1, 2])) {
    return(
      as.numeric(
        m1[1, 2]
      )
    )
  }

  m2 <- str_match(
    x2,
    regex(
      "([0-9]+)\\s*humans?",
      ignore_case = TRUE
    )
  )

  if (!is.na(m2[1, 2])) {
    return(
      as.numeric(
        m2[1, 2]
      )
    )
  }

  nums <- str_extract_all(
    x2,
    "[0-9]+"
  )[[1]]

  if (length(nums) == 0) {
    return(NA_real_)
  }

  as.numeric(
    nums[1]
  )
}


categorise_sample_size <- function(n) {

  case_when(

    is.na(n) ~
      NA_character_,

    n < 1000 ~
      "<1,000",

    n >= 1000 &
      n < 20000 ~
      "1,000-19,999",

    n >= 20000 &
      n < 100000 ~
      "20,000-99,999",

    n >= 100000 ~
      ">=100,000",

    TRUE ~ NA_character_
  )
}


clean_setting <- function(x) {

  x <- clean_text(x)

  case_when(

    str_detect(
      x,
      regex(
        paste0(
          "mixed rural.*urban|rural.*urban|",
          "general population \\(rural.?&.?urban\\)"
        ),
        ignore_case = TRUE
      )
    ) ~ "Mixed rural and urban",

    str_detect(
      x,
      regex(
        "^urban|urban population",
        ignore_case = TRUE
      )
    ) ~ "Urban",

    str_detect(
      x,
      regex(
        "^rural|rural population",
        ignore_case = TRUE
      )
    ) ~ "Rural",

    TRUE ~ "Other/unclear"
  )
}


categorise_decade <- function(year) {

  case_when(

    is.na(year) ~
      NA_character_,

    year < 1980 ~
      "<=1970s",

    year >= 1980 &
      year < 1990 ~
      "1980s",

    year >= 1990 &
      year < 2000 ~
      "1990s",

    year >= 2000 &
      year < 2010 ~
      "2000s",

    year >= 2010 &
      year < 2020 ~
      "2010s",

    year >= 2020 ~
      "2020s",

    TRUE ~ NA_character_
  )
}


categorise_status <- function(
    end_year
) {

  case_when(

    is.na(end_year) ~
      "Ongoing",

    TRUE ~
      "Ended"
  )
}


parse_followup_months <- function(x) {

  x <- tolower(
    clean_text(x)
  )

  if (
    is.na(x) ||
      x == ""
  ) {
    return(NA_real_)
  }

  if (
    str_detect(
      x,
      paste0(
        "baseline|single time point|point-in-time|",
        "single cross-sectional|cross-sectional survey"
      )
    )
  ) {
    return(0)
  }

  if (
    str_detect(
      x,
      "days?\\s*3\\s*&\\s*7"
    ) ||
      str_detect(
        x,
        "days?\\s*2\\s*&\\s*7"
      )
  ) {
    return(14 / 30.4375)
  }

  if (
    str_detect(
      x,
      "14 days|2 weeks"
    )
  ) {
    return(14 / 30.4375)
  }

  if (
    str_detect(
      x,
      "9 days"
    )
  ) {
    return(9 / 30.4375)
  }

  if (
    str_detect(
      x,
      "6 weeks"
    )
  ) {
    return(6 / 4.345)
  }

  if (
    str_detect(
      x,
      "weekly"
    )
  ) {
    return(7 / 30.4375)
  }

  if (
    str_detect(
      x,
      "bi weekly|bi-weekly"
    )
  ) {
    return(14 / 30.4375)
  }

  if (
    str_detect(
      x,
      "hourly"
    )
  ) {
    return(
      1 / 24 / 30.4375
    )
  }

  if (
    str_detect(
      x,
      "daily & monthly"
    )
  ) {
    return(1)
  }

  if (
    str_detect(
      x,
      "daily"
    )
  ) {
    return(
      1 / 30.4375
    )
  }

  if (
    str_detect(
      x,
      "anual|annual"
    )
  ) {
    return(12)
  }

  if (
    str_detect(
      x,
      "12-month"
    )
  ) {
    return(12)
  }

  if (
    str_detect(
      x,
      "24-month"
    )
  ) {
    return(24)
  }

  if (
    str_detect(
      x,
      paste0(
        "continuous|periodic|event-driven|",
        "longitudinal|registration|surveillance"
      )
    )
  ) {

    nums <- suppressWarnings(
      as.numeric(
        str_extract_all(
          x,
          "\\d+\\.?\\d*"
        )[[1]]
      )
    )

    nums <- nums[
      !is.na(nums)
    ]

    if (
      length(nums) > 0
    ) {
      return(
        max(nums)
      )
    }

    return(
      NA_real_
    )
  }

  if (
    str_detect(
      x,
      "^\\d+\\.?\\d*$"
    )
  ) {
    return(
      as.numeric(x)
    )
  }

  if (
    str_detect(
      x,
      "week"
    )
  ) {

    num <- suppressWarnings(
      parse_number(x)
    )

    if (
      !is.na(num)
    ) {
      return(
        num / 4.345
      )
    }
  }

  if (
    str_detect(
      x,
      "day"
    )
  ) {

    num <- suppressWarnings(
      parse_number(x)
    )

    if (
      !is.na(num)
    ) {
      return(
        num / 30.4375
      )
    }
  }

  if (
    str_detect(
      x,
      "month"
    )
  ) {

    num <- suppressWarnings(
      parse_number(x)
    )

    if (
      !is.na(num)
    ) {
      return(num)
    }
  }

  num <- suppressWarnings(
    parse_number(x)
  )

  if (
    !is.na(num)
  ) {
    return(num)
  }

  NA_real_
}


categorise_followup_range <- function(
    min_m,
    max_m
) {

  case_when(

    is.na(min_m) &
      is.na(max_m) ~
      NA_character_,

    !is.na(max_m) &
      max_m == 0 ~
      "Baseline only",

    !is.na(max_m) &
      max_m > 0 &
      max_m <= 1 ~
      ">0-1 month",

    !is.na(max_m) &
      max_m > 1 &
      max_m <= 6 ~
      ">1-6 months",

    !is.na(max_m) &
      max_m > 6 &
      max_m <= 12 ~
      ">6-12 months",

    !is.na(max_m) &
      max_m > 12 &
      max_m <= 24 ~
      ">12-24 months",

    !is.na(max_m) &
      max_m > 24 ~
      ">24 months",

    TRUE ~
      NA_character_
  )
}


summ_cat <- function(
    data,
    var
) {

  tmp <- data %>%

    filter(

      !is.na(
        {{ var }}
      ),

      {{ var }} != "",

      !str_detect(
        as.character(
          {{ var }}
        ),
        regex(
          "unclear|other",
          ignore_case = TRUE
        )
      )
    )

  denom <- nrow(tmp)

  tmp %>%

    count(
      {{ var }},
      name = "n"
    ) %>%

    mutate(
      percent =
        round(
          100 * n / denom,
          1
        )
    ) %>%

    rename(
      category =
        {{ var }}
    ) %>%

    mutate(
      category =
        as.character(category)
    )
}


summ_binary_vars <- function(
    data,
    vars,
    labels,
    variable_name
) {

  tibble(

    variable =
      variable_name,

    category =
      labels,

    n =
      map_int(
        vars,
        ~ sum(
          is_yes(
            data[[.x]]
          ),
          na.rm = TRUE
        )
      ),

    percent =
      round(
        100 * n /
          nrow(data),
        1
      )
  ) %>%

    filter(
      n > 0
    ) %>%

    arrange(
      desc(n),
      category
    )
}
```

# 5. Harmonise cohort names

```{r cohort-name-harmonisation}
name_crosswalk <- tibble::tribble(

  ~raw_name,
  ~harmonized_name,

  "Manyatta HDSS",
  "Manyatta Karemo HDSS",

  "Karemo HDSS",
  "Manyatta Karemo HDSS",

  "Manyatta/Karemo HDSS",
  "Manyatta Karemo HDSS",

  "Kilifi HDSS2",
  "Kilifi HDSS",

  "Kisumu Ahero/KEMRI CDC",
  "Kisumu KEMRI CDC",

  "Dinamo HDSS/Dikgale",
  "Dinamo HDSS",

  "MLSFH/gen-malawi",
  "MLSFH",

  "Djicoroni Para&Banconi, Bamako_CHAMPS_Network",
  "Harar_Djicoroni_CHAMPS",

  "Harar_CHAMPS_Network",
  "Harar_Djicoroni_CHAMPS",

  paste0(
    "GRT-INSPIRED HDSS (Gauteng Research Triangle Initiative ",
    "for the Study of Population, Infrastructure and Regional ",
    "Economic Development)"
  ),
  "GRT-INSPIRED HDSS",

  "GRT-INSPIRED HDSS (Gauteng)",
  "GRT-INSPIRED HDSS",

  paste0(
    "C-SHARP HDSS (Cape Town Surveillance through ",
    "Healthcare Action Research Project)"
  ),
  "C-SHARP HDSS",

  "C-SHARP HDSS (Western Cape)",
  "C-SHARP HDSS",

  "BAMMISHO",
  "BAMMISHO HDSS",

  paste0(
    "Northwestern Tigray Rabies Exposure Cohort ",
    "(Suhul Hospital Catchment)"
  ),
  "Northwestern Tigray Rabies Exposure Cohort",

  paste0(
    "PATHOME Study (Pathogen Transmission and Health ",
    "Outcome Models of Enteric Disease)"
  ),
  "PATHOME Study",

  "Dielmo & Ndiop Malaria Cohort",
  "Dielmo and Ndiop Malaria Cohorts (Senegal)"
)


harmonize_cohort_name <- function(
    x,
    crosswalk = name_crosswalk
) {

  x_clean <- clean_text(x)

  matched <-
    crosswalk$harmonized_name[
      match(
        x_clean,
        crosswalk$raw_name
      )
    ]

  ifelse(
    is.na(matched),
    x_clean,
    matched
  )
}


# Preserve the original cohort/site record
# before harmonising cohort names.

cohort_info <- cohort_info %>%

  mutate(

    cohort_record_id =
      row_number(),

    cohort_name_raw =
      clean_text(
        cohort_name
      ),

    cohort_name =
      harmonize_cohort_name(
        cohort_name_raw
      )
  )


oh_outcomes <- oh_outcomes %>%

  mutate(

    cohort_name_raw =
      clean_text(
        cohort_name
      ),

    cohort_name =
      harmonize_cohort_name(
        cohort_name_raw
      )
  )


oh_exposures <- oh_exposures %>%

  mutate(

    cohort_name_raw =
      clean_text(
        cohort_name
      ),

    cohort_name =
      harmonize_cohort_name(
        cohort_name_raw
      )
  )


outcome_details <- outcome_details %>%

  mutate(

    cohort_name_raw =
      clean_text(
        cohort_name
      ),

    cohort_name =
      harmonize_cohort_name(
        cohort_name_raw
      )
  )
```

# 6. Clean and collapse outcome data

```{r clean-outcomes}
if (
  all(
    c(
      "Brucellosis...30",
      "Brucellosis...36"
    ) %in%
      names(oh_outcomes)
  )
) {

  oh_outcomes <-
    oh_outcomes %>%

    mutate(

      Brucellosis =
        ifelse(

          to_yesno_strict(
            `Brucellosis...30`
          ) == "Yes" |

            to_yesno_strict(
              `Brucellosis...36`
            ) == "Yes",

          "Yes",
          "No"
        )
    ) %>%

    select(
      -`Brucellosis...30`,
      -`Brucellosis...36`
    )
}


outcome_cols_raw <-
  setdiff(

    names(
      oh_outcomes
    ),

    c(
      "cohort_name",
      "cohort_name_raw"
    )
  )


oh_outcomes_cohort <-
  oh_outcomes %>%

  mutate(
    across(
      all_of(
        outcome_cols_raw
      ),
      to_yesno_strict
    )
  ) %>%

  group_by(
    cohort_name
  ) %>%

  summarise(

    across(

      all_of(
        outcome_cols_raw
      ),

      ~ ifelse(
        any(
          .x == "Yes"
        ),
        "Yes",
        "No"
      )
    ),

    .groups =
      "drop"
  )


outcome_list_by_cohort <-
  oh_outcomes_cohort %>%

  rowwise() %>%

  mutate(

    outcomes_list =
      paste(

        names(
          c_across(
            all_of(
              outcome_cols_raw
            )
          )
        )[
          c_across(
            all_of(
              outcome_cols_raw
            )
          ) == "Yes"
        ],

        collapse =
          "; "
      ),

    outcomes_list =
      ifelse(
        outcomes_list == "",
        NA_character_,
        outcomes_list
      )
  ) %>%

  ungroup() %>%

  select(
    cohort_name,
    outcomes_list
  )
```

# 7. Define and collapse One Health exposure variables

```{r exposure-groups}
animal_vars <- c(

  "Wildlife/wildbird Proximity or contact",
  "Livestock ownership (general)",
  "Livestock ownership (Pig)",
  "Poultry/birds Ownership",
  "Livestock ownership (Goat)",
  "Livestock ownership (Cattle)",
  "Livestock ownership (Sheep)",
  "Small animals",
  "Bats contact",
  "Bushmeat consumption",
  "Cat contact/ownership",
  "Dog contact/bites",
  "Animal Sampling",
  "Animal Health Practices(Surveys - Biosecurity, vaccination)",
  "Antibiotic use in animals",
  "Dog Vaccination"
)


ecosystem_vars <- c(

  "WASH (Hand washing, saniation, access to clean water)",
  "Natural water body contact",
  "Water waste management",
  "Solid waste management",
  "WASH(Toilet type)",
  "Food secutity (nutrition)",
  "Air pollution(Particulate Matter (PM2.5 or other)",
  "Air pollution(Energy Source Cooking)",
  "Air Pollution(Dust exposure, proximity to Road etc)",
  "Seasonality",
  "Climate Variables(Temp.)",
  "Climate Variables(Humidity)",
  "Climate Variables(Dryness)",
  "Climate Variables(Rainfall)",
  "Greenness",
  "Noise Pollution",
  "Chemical exposure",
  "Flooding context",
  "Soil contact / geophagy",
  "Other_Environmental sampling"
)


vector_vars <- c(

  "Use of bed nets",
  "Entomological survey - Mosquitoes(densities etc)",
  "Entomological survey - Ticks",
  "Vector exposure(mosquitoes, snails, ticks)"
)


socioeconomic_vars <- c(
  "Socio-economic Indicators"
)


other_contextual_vars <- c(

  "Geolocation (GPS)",
  "Residency(Rural/Urban)"
)


occupation_animal_pattern <-
  paste(

    c(
      "pastoral",
      "poultry",
      "livestock",
      "hunting",
      "hunting and farming",
      "farming and pastoralism"
    ),

    collapse =
      "|"
  )


occupation_ecosystem_pattern <-
  paste(

    c(
      "farming",
      "agriculture",
      "fishermen",
      "fishing",
      "mining",
      "pastoral",
      "hunting",
      "poultry"
    ),

    collapse =
      "|"
  )


oh_exposures2 <-
  oh_exposures %>%

  mutate(

    occupation_value =
      str_to_lower(
        clean_text(
          `Occupation_ OH related (Specify)`
        )
      ),

    occupation_animal_related =
      ifelse(
        str_detect(
          occupation_value,
          occupation_animal_pattern
        ),
        "Yes",
        "No"
      ),

    occupation_ecosystem_related =
      ifelse(
        str_detect(
          occupation_value,
          occupation_ecosystem_pattern
        ),
        "Yes",
        "No"
      )
  )


animal_cols <-
  unique(
    c(
      intersect(
        animal_vars,
        names(oh_exposures2)
      ),
      "occupation_animal_related"
    )
  )


ecosystem_cols <-
  unique(
    c(
      intersect(
        ecosystem_vars,
        names(oh_exposures2)
      ),
      "occupation_ecosystem_related"
    )
  )


vector_cols <-
  intersect(
    vector_vars,
    names(oh_exposures2)
  )


socioeconomic_cols <-
  intersect(
    socioeconomic_vars,
    names(oh_exposures2)
  )


other_contextual_cols <-
  intersect(
    other_contextual_vars,
    names(oh_exposures2)
  )


all_exposure_cols <-
  unique(
    c(
      animal_cols,
      ecosystem_cols,
      vector_cols,
      socioeconomic_cols,
      other_contextual_cols
    )
  )


oh_exposures_cohort <-
  oh_exposures2 %>%

  mutate(
    across(
      all_of(
        all_exposure_cols
      ),
      to_yesno_exposure
    )
  ) %>%

  group_by(
    cohort_name
  ) %>%

  summarise(

    across(

      all_of(
        all_exposure_cols
      ),

      ~ ifelse(
        any(
          .x == "Yes"
        ),
        "Yes",
        "No"
      )
    ),

    .groups =
      "drop"
  )


cohort_domain_flags <-
  oh_exposures_cohort %>%

  mutate(

    human_outcome =
      "Yes",

    animal_exposure_domain =
      ifelse(
        yes_count(
          .,
          animal_cols
        ) > 0,
        "Yes",
        "No"
      ),

    ecosystem_exposure_domain =
      ifelse(
        yes_count(
          .,
          ecosystem_cols
        ) > 0,
        "Yes",
        "No"
      ),

    vector_exposure_domain =
      ifelse(
        yes_count(
          .,
          vector_cols
        ) > 0,
        "Yes",
        "No"
      ),

    socioeconomic_exposure_domain =
      ifelse(
        yes_count(
          .,
          socioeconomic_cols
        ) > 0,
        "Yes",
        "No"
      ),

    other_contextual_exposure_domain =
      ifelse(
        yes_count(
          .,
          other_contextual_cols
        ) > 0,
        "Yes",
        "No"
      )
  ) %>%

  select(
    cohort_name,
    human_outcome,
    animal_exposure_domain,
    ecosystem_exposure_domain,
    vector_exposure_domain,
    socioeconomic_exposure_domain,
    other_contextual_exposure_domain
  )
```

# 8. Outcome-detail, population, method, assay and biospecimen flags

```{r outcome-detail-flags}
outcome_details_cohort <-
  outcome_details %>%

  group_by(
    cohort_name
  ) %>%

  summarise(

    population_type_text =
      collapse_unique_text(
        `Population Type/setting for specific outcome`
      ),

    assessment_methods_text =
      collapse_unique_text(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`
      ),

    biospecimens_text =
      collapse_unique_text(
        Biospecimen
      ),

    .groups =
      "drop"
  )


population_flags <-
  outcome_details %>%

  group_by(
    cohort_name
  ) %>%

  summarise(

    age_children =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "child|children|infant|newborn|neonat|under 5|<5|0–5|0-5"
      ),

    age_adolescents =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "adolescent|youth|teen"
      ),

    age_adults =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "adult|men and women|general population|aged ≥|above 40"
      ),

    age_all_ages =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "all ages"
      ),

    sex_male =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "\\bmen\\b|male"
      ),

    sex_female =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "\\bwomen\\b|female|mother"
      ),

    sex_both =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "both sexes|all ages & genders|men and women|all ages"
      ),

    special_pregnant =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "pregnan|gravid|antenatal|anc"
      ),

    special_hiv =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "hiv|art patients|people living with hiv"
      ),

    special_hospital =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "hospital|in-patients|facility"
      ),

    special_deceased =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "deceased|all deaths|mortality"
      ),

    special_animals =
      flag_detect(
        `Population Type/setting for specific outcome`,
        "cattle|goats|sheep|dogs|livestock|animals|ruminants|equines"
      ),

    .groups =
      "drop"
  )


method_flags <-
  outcome_details %>%

  group_by(
    cohort_name
  ) %>%

  summarise(

    questionnaire_data =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        paste0(
          "questionnaire|survey|interview|self-report|caregiver report|",
          "structured questionnaire|respondent report|oral history|",
          "symptom diary|self-administered|caregiver recall"
        )
      ),

    clinical_data =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        paste0(
          "clinical|physical examination|clinical examination|blood pressure|",
          "spirometry|anthropometr|clinical measurement|medical records|",
          "radiology|clinical assessment|clinical follow-up|clinical screening|",
          "clinical exams|clinical diagnosis|clinical evaluation|",
          "physician examination|delivery room measurements|",
          "physical exam|standardized physical measurement|physical measurements"
        )
      ),

    routine_data =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        paste0(
          "routine|registry|record|hdss|surveillance|census|",
          "demographic surveillance|verbal autopsy|facility data|",
          "data linkage|vital status records|household rosters|",
          "community reporting|reporting|enumeration|",
          "mortality records|clinic data"
        )
      ),

    laboratory_data =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        paste0(
          "laboratory|serology|pcr|qpcr|rt-pcr|elisa|microscopy|culture|",
          "sequencing|assay|genexpert|hematology|blood count|biomarker|",
          "molecular|antibody|antigen|neutralization|haemoglobin|hemoglobin|",
          "diagnostics|kato-katz|urine filtration|blood film|",
          "microbiology|microbiological"
        )
      ),

    case_based_passive_sentinel_detection =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        paste0(
          "review of hospital diagnosis records|hospital-based morbidity surveillance|",
          "facility-based morbidity surveillance|hospital-based surveillance|",
          "facility-based surveillance|",
          "hospital-based surveillance and laboratory confirmation|",
          "facility-based surveillance and laboratory confirmation|",
          "health-facility based surveillance|health facility diagnosis|",
          "clinical examination and diagnosis at health facilities|",
          "passive case detection|passive surveillance|",
          "morbidity surveillance|morbidity reporting|",
          "clinic follow-up|vital status records, clinic follow-up|sentinel"
        )
      ),

    active_surveillance_detection_raw =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        paste0(
          "household surveillance|household-based surveillance|household visits|",
          "active case-finding|active case detection|active weekly household visits|",
          "active home visits|repeated household clinical symptom surveillance|",
          "household surveillance interviews|community follow-up|",
          "continuous surveillance|longitudinal surveillance|",
          "routine hdss surveillance|follow-up tracing|surveillance surveys|",
          "demographic surveillance|planned and nested studies|",
          "household follow-up interviews|weekly household visits"
        )
      ),

    .groups =
      "drop"
  ) %>%

  mutate(

    active_surveillance_detection =
      ifelse(
        case_based_passive_sentinel_detection ==
          "Yes",
        "No",
        "Yes"
      )
  )


assay_flags <-
  outcome_details %>%

  group_by(
    cohort_name
  ) %>%

  summarise(

    assay_serology =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "serology|antibody|igg|igm|neutralization|microneutralization|hemagglutination"
      ),

    assay_pcr =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "\\bpcr\\b|qpcr|rt-pcr|genexpert|molecular"
      ),

    assay_microscopy =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "microscopy|blood smear|kato-katz|urine filtration"
      ),

    assay_culture =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "culture|bactec|microbiological"
      ),

    assay_elisa =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "elisa"
      ),

    assay_sequencing =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "sequencing|genotyping|genome-wide|methylation|beadchip|pyrosequencing"
      ),

    assay_hematology =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "hemoglobin|haemoglobin|full blood count|complete blood count|hemocue"
      ),

    assay_biomarkers =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "biomarker|cholesterol|creatinine|crp|glucose|hba1c"
      ),

    assay_anthropometry =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "anthropometr|height|weight|bmi|muac|z-score"
      ),

    assay_verbal_autopsy =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "verbal autopsy|social autopsy"
      ),

    assay_questionnaire =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "questionnaire|survey|interview"
      ),

    assay_rdt =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "rapid diagnostic|rapid test|\\brdt\\b"
      ),

    assay_imaging =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "radiology|x-ray"
      ),

    assay_spirometry =
      flag_detect(
        `Outcome_assessment_method(specific laboratory methods_or_survey)`,
        "spirometry|lung function"
      ),

    .groups =
      "drop"
  )


biospecimen_flags <-
  outcome_details %>%

  group_by(
    cohort_name
  ) %>%

  summarise(

    sample_blood =
      flag_detect(
        Biospecimen,
        "blood|whole blood|capillary blood|venous blood|finger-prick"
      ),

    sample_serum =
      flag_detect(
        Biospecimen,
        "serum"
      ),

    sample_plasma =
      flag_detect(
        Biospecimen,
        "plasma"
      ),

    sample_dbs =
      flag_detect(
        Biospecimen,
        "dried blood spot"
      ),

    sample_dna =
      flag_detect(
        Biospecimen,
        "\\bdna\\b"
      ),

    sample_urine =
      flag_detect(
        Biospecimen,
        "urine"
      ),

    sample_stool =
      flag_detect(
        Biospecimen,
        "stool|faeces|feces"
      ),

    sample_sputum =
      flag_detect(
        Biospecimen,
        "sputum|gastric aspirate"
      ),

    sample_swab =
      flag_detect(
        Biospecimen,
        paste0(
          "swab|nasal swab|nasopharyngeal|",
          "oropharyngeal|rectal swab|vaginal swab"
        )
      ),

    sample_tissue =
      flag_detect(
        Biospecimen,
        "tissue|biops"
      ),

    sample_placenta =
      flag_detect(
        Biospecimen,
        "placenta|cord blood|umbilical cord"
      ),

    sample_saliva =
      flag_detect(
        Biospecimen,
        "saliva"
      ),

    sample_hair =
      flag_detect(
        Biospecimen,
        "hair"
      ),

    sample_mosquito =
      flag_detect(
        Biospecimen,
        "mosquito"
      ),

    sample_animal =
      flag_detect(
        Biospecimen,
        "animal|brain tissue|animal feces"
      ),

    sample_environmental =
      flag_detect(
        Biospecimen,
        "water|soil|surface|food|hand rinse|toy rinse|air"
      ),

    .groups =
      "drop"
  )


human_sampling_flags <-
  biospecimen_flags %>%

  mutate(

    human_sampling =
      ifelse(

        sample_blood == "Yes" |
          sample_serum == "Yes" |
          sample_plasma == "Yes" |
          sample_dbs == "Yes" |
          sample_dna == "Yes" |
          sample_urine == "Yes" |
          sample_stool == "Yes" |
          sample_sputum == "Yes" |
          sample_swab == "Yes" |
          sample_tissue == "Yes" |
          sample_placenta == "Yes" |
          sample_saliva == "Yes" |
          sample_hair == "Yes",

        "Yes",
        "No"
      )
  ) %>%

  select(
    cohort_name,
    human_sampling
  )
```

# 9. Build the cohort-level master dataset

```{r build-cohort-master}
cohort_info_clean <-
  cohort_info %>%

  transmute(

    cohort_record_id,

    cohort_name,

    cohort_name_raw,

    country =
      clean_text(
        Country
      ),

    start_year =
      `Start Year`,

    end_year =
      `End Year`,

    recruitment_method =
      clean_text(
        `Recruitment Method`
      ),

    contact_details =
      clean_text(
        `Contact details`
      ),

    cohort_link =
      clean_text(
        `Link to the cohort`
      ),

    participant_size =
      `Sample Size(participants) reported in the cohort website`,

    setting =
      clean_text(
        Setting
      ),

    data_sharing_potential =
      clean_text(
        `Data Sharing Potential`
      ),

    clinical_data_link =
      clean_text(
        `Link to Clinical data`
      ),

    open_future_collaboration =
      clean_text(
        `Open for future questions/collaborations`
      ),

    biobank_samples =
      clean_text(
        `Biobank of Samples`
      ),

    hdss =
      clean_text(
        `HDSS?`
      ),

    mention_one_health =
      clean_text(
        `Explicitly mention any One Health Framework?`
      )
  )


# cohort_info is the reference cohort/site list.
# left_join prevents additional cohort records
# from being created during merging.

cohort_master <-
  cohort_info_clean %>%

  left_join(
    cohort_domain_flags,
    by = "cohort_name"
  ) %>%

  left_join(
    oh_exposures_cohort,
    by = "cohort_name"
  ) %>%

  left_join(
    oh_outcomes_cohort,
    by = "cohort_name"
  ) %>%

  left_join(
    outcome_list_by_cohort,
    by = "cohort_name"
  ) %>%

  left_join(
    outcome_details_cohort,
    by = "cohort_name"
  ) %>%

  left_join(
    population_flags,
    by = "cohort_name"
  ) %>%

  left_join(
    method_flags,
    by = "cohort_name"
  ) %>%

  left_join(
    assay_flags,
    by = "cohort_name"
  ) %>%

  left_join(
    biospecimen_flags,
    by = "cohort_name"
  ) %>%

  left_join(
    human_sampling_flags,
    by = "cohort_name"
  )


if (
  nrow(cohort_master) !=
    nrow(cohort_info_clean)
) {

  stop(
    paste0(
      "The merge changed the number of cohort/site records. ",
      "Check duplicate rows in cohort-level lookup tables."
    )
  )
}


binary_cols <-
  unique(
    c(

      "human_outcome",
      "animal_exposure_domain",
      "ecosystem_exposure_domain",
      "vector_exposure_domain",
      "socioeconomic_exposure_domain",
      "other_contextual_exposure_domain",

      "human_sampling",

      "questionnaire_data",
      "clinical_data",
      "routine_data",
      "laboratory_data",

      "case_based_passive_sentinel_detection",
      "active_surveillance_detection",

      "age_children",
      "age_adolescents",
      "age_adults",
      "age_all_ages",

      "sex_male",
      "sex_female",
      "sex_both",

      "special_pregnant",
      "special_hiv",
      "special_hospital",
      "special_deceased",
      "special_animals",

      "assay_serology",
      "assay_pcr",
      "assay_microscopy",
      "assay_culture",
      "assay_elisa",
      "assay_sequencing",
      "assay_hematology",
      "assay_biomarkers",
      "assay_anthropometry",
      "assay_verbal_autopsy",
      "assay_questionnaire",
      "assay_rdt",
      "assay_imaging",
      "assay_spirometry",

      "sample_blood",
      "sample_serum",
      "sample_plasma",
      "sample_dbs",
      "sample_dna",
      "sample_urine",
      "sample_stool",
      "sample_sputum",
      "sample_swab",
      "sample_tissue",
      "sample_placenta",
      "sample_saliva",
      "sample_hair",
      "sample_mosquito",
      "sample_animal",
      "sample_environmental",

      outcome_cols_raw,

      all_exposure_cols
    )
  )


binary_cols <-
  binary_cols[
    binary_cols %in%
      names(cohort_master)
  ]


cohort_master <-
  cohort_master %>%

  mutate(
    across(
      all_of(
        binary_cols
      ),
      force_yesno
    )
  ) %>%

  mutate(
    human_outcome =
      "Yes"
  )
```

# 10. Construct cleaned One Health subdomains and domains

```{r cleaned-one-health-domains}
human_behaviour_vars <- c(

  "Use of bed nets",
  "Bushmeat consumption",
  "Natural water body contact",
  "Soil contact / geophagy",
  "Cat contact/ownership",
  "Dog contact/bites"
)


human_contextual_vars <- c(

  "Socio-economic Indicators",
  "occupation_animal_related",
  "occupation_ecosystem_related",
  "Residency(Rural/Urban)",
  "Geolocation (GPS)"
)


animal_contact_vars <- c(

  "Wildlife/wildbird Proximity or contact",
  "Livestock ownership (general)",
  "Livestock ownership (Pig)",
  "Poultry/birds Ownership",
  "Livestock ownership (Goat)",
  "Livestock ownership (Cattle)",
  "Livestock ownership (Sheep)",
  "Small animals",
  "Bats contact",
  "Cat contact/ownership",
  "Dog contact/bites"
)


animal_health_vars <- c(

  "Animal Health Practices(Surveys - Biosecurity, vaccination)",
  "Antibiotic use in animals",
  "Dog Vaccination"
)


vector_entomological_vars <- c(

  "Entomological survey - Mosquitoes(densities etc)",
  "Entomological survey - Ticks"
)


environmental_monitoring_vars <- c(

  "Air pollution(Particulate Matter (PM2.5 or other)",
  "Climate Variables(Temp.)",
  "Climate Variables(Humidity)",
  "Climate Variables(Dryness)",
  "Climate Variables(Rainfall)"
)


environmental_exposure_vars <- c(

  "Air pollution(Energy Source Cooking)",
  "Air Pollution(Dust exposure, proximity to Road etc)",
  "Noise Pollution",
  "Chemical exposure",
  "Flooding context",
  "Natural water body contact",
  "Water waste management",
  "Solid waste management",
  "Soil contact / geophagy",
  "Greenness",
  "WASH (Hand washing, saniation, access to clean water)",
  "WASH(Toilet type)",
  "Food secutity (nutrition)",
  "Seasonality"
)


climate_vars <- c(

  "Climate Variables(Temp.)",
  "Climate Variables(Humidity)",
  "Climate Variables(Dryness)",
  "Climate Variables(Rainfall)"
)


cohort_master <-
  cohort_master %>%

  mutate(

    human_sampling_cleaned =
      ifelse(
        human_sampling ==
          "Yes",
        "Yes",
        "No"
      ),

    human_behaviour_exposure_cleaned =
      ifelse(
        yes_count(
          .,
          human_behaviour_vars
        ) > 0,
        "Yes",
        "No"
      ),

    human_contextual_cleaned =
      ifelse(
        yes_count(
          .,
          human_contextual_vars
        ) > 0,
        "Yes",
        "No"
      ),

    animal_contact_cleaned =
      ifelse(
        yes_count(
          .,
          animal_contact_vars
        ) > 0,
        "Yes",
        "No"
      ),

    animal_sampling_cleaned =
      ifelse(
        yes_count(
          .,
          c(
            "Animal Sampling",
            "sample_animal"
          )
        ) > 0,
        "Yes",
        "No"
      ),

    vector_entomological_cleaned =
      ifelse(
        yes_count(
          .,
          c(
            vector_entomological_vars,
            "sample_mosquito"
          )
        ) > 0,
        "Yes",
        "No"
      ),

    animal_health_practices_cleaned =
      ifelse(
        yes_count(
          .,
          animal_health_vars
        ) > 0,
        "Yes",
        "No"
      ),

    environmental_sampling_cleaned =
      ifelse(
        yes_count(
          .,
          c(
            "Other_Environmental sampling",
            "sample_environmental"
          )
        ) > 0,
        "Yes",
        "No"
      ),

    environmental_monitoring_cleaned =
      ifelse(
        yes_count(
          .,
          environmental_monitoring_vars
        ) > 0,
        "Yes",
        "No"
      ),

    environmental_exposure_assessment_cleaned =
      ifelse(
        yes_count(
          .,
          environmental_exposure_vars
        ) > 0,
        "Yes",
        "No"
      ),

    climate_data_cleaned =
      ifelse(
        yes_count(
          .,
          climate_vars
        ) > 0,
        "Yes",
        "No"
      ),

    human_domain_cleaned =
      ifelse(

        human_sampling_cleaned ==
          "Yes" |

          human_behaviour_exposure_cleaned ==
            "Yes" |

          human_contextual_cleaned ==
            "Yes" |

          human_outcome ==
            "Yes",

        "Yes",
        "No"
      ),

    animal_domain_cleaned =
      ifelse(

        animal_contact_cleaned ==
          "Yes" |

          animal_sampling_cleaned ==
            "Yes" |

          vector_entomological_cleaned ==
            "Yes" |

          animal_health_practices_cleaned ==
            "Yes",

        "Yes",
        "No"
      ),

    environmental_domain_cleaned =
      ifelse(

        environmental_sampling_cleaned ==
          "Yes" |

          environmental_monitoring_cleaned ==
            "Yes" |

          environmental_exposure_assessment_cleaned ==
            "Yes" |

          climate_data_cleaned ==
            "Yes" |

          human_contextual_cleaned ==
            "Yes",

        "Yes",
        "No"
      )
  )


# Final exclusion used in the review.

cohort_master <-
  cohort_master %>%

  filter(
    cohort_name !=
      "SIBS Genomic HDSS"
  ) %>%

  arrange(
    cohort_record_id
  ) %>%

  mutate(

    cohort_id =
      row_number(),

    broad_interface_cleaned =
      case_when(

        human_domain_cleaned ==
          "Yes" &

          animal_domain_cleaned ==
            "Yes" &

          environmental_domain_cleaned ==
            "Yes" ~

          "Human-Animal-Environmental",

        human_domain_cleaned ==
          "Yes" &

          animal_domain_cleaned ==
            "No" &

          environmental_domain_cleaned ==
            "Yes" ~

          "Human-Environmental",

        human_domain_cleaned ==
          "Yes" &

          animal_domain_cleaned ==
            "Yes" &

          environmental_domain_cleaned ==
            "No" ~

          "Human-Animal",

        human_domain_cleaned ==
          "Yes" &

          animal_domain_cleaned ==
            "No" &

          environmental_domain_cleaned ==
            "No" ~

          "Human only",

        TRUE ~
          NA_character_
      )
  )


cat(
  "Final cohort records:",
  nrow(cohort_master),
  "\n"
)


cat(
  "Unique cohort IDs:",
  n_distinct(
    cohort_master$cohort_id
  ),
  "\n"
)


cat(
  "Unique harmonised cohort names:",
  n_distinct(
    cohort_master$cohort_name
  ),
  "\n"
)


repeated_cohort_names <-
  cohort_master %>%

  count(
    cohort_name,
    sort = TRUE
  ) %>%

  filter(
    n > 1
  )


repeated_cohort_names


if (
  nrow(cohort_master) != 174
) {

  warning(
    paste0(
      "Expected 174 cohort/site records, but found ",
      nrow(cohort_master),
      ". Check the source data and exclusions."
    )
  )
}


write.csv(
  cohort_master,
  "cohort_master_dataset.csv",
  row.names = FALSE
)
```

# 11. Outcome classification

```{r outcome-classification}
outcome_cols <- c(

  "Malaria",
  "Tuberculosis",
  "HIV/AIDS",
  "Vaccination Coverage",
  "Demographic trends(births, migrations, pregnancies etc)",
  "Mortality trends & Causes",
  "Diarrhea",
  "Maternal, fertility & Child Health",
  "Nutrition",
  "Measles",
  "Chickenpox",
  "Rotavirus",
  "Anaemia",
  "S. pneumoniae",
  "Drug Resistance",
  "Cardiovascular diseases",
  "Cancers",
  "Chronic respiratory diseases",
  "Diabetes",
  "Mental and neurological disorders",
  "Musculoskeletal conditions",
  "Chronic kidney disease",
  "Acute Respiratory illnesses(ARI)",
  "Influenza Like Illnesses (ILI)",
  "Acute Febrile Illnesses (AFI)",
  "Schistosomiasis",
  "Filariasis (specifically Lymphatic Filariasis)",
  "Rabies",
  "Seasonal Influenza (Flu virus)",
  "COVID-19 (SARS-CoV-2)",
  "Respiratory Syncytial Virus (RSV)",
  "Parainfluenza Virus",
  "MERS-CoV",
  "Leptospirosis",
  "Eye health and vision",
  "Anthrax",
  "Plague (Yersinia pestis)",
  "Toxoplasmosis",
  "Borrelia",
  "Bovine tuberculosis (Mycobacterium bovis)",
  "Campylobacteriosis",
  "Listeriosis",
  "Salmonellosis",
  "Yersiniosis",
  "Avian influenza (also respiratory but zoonotic)",
  "Dengue Fever",
  "Chikungunya",
  "Zika Virus",
  "Yellow Fever",
  "West Nile Virus",
  "Rift Valley Fever",
  "Human African Trypanosomiasis (Sleeping Sickness)",
  "Leishmaniasis",
  "Tick-borne Encephalitis",
  "Lyme Disease",
  "Crimean-Congo Hemorrhagic Fever",
  "Pertusis",
  "Dracunculus medinensis",
  "Acute Illnesses(general)",
  "soil-transmitted helminths.",
  "NCDS(General mentioned)",
  "Hepatitis",
  "Hypertension",
  "Clinical Examination(blood pressure, anthropometry)",
  "dental examination",
  "Menengitis",
  "Fasciolosis",
  "Injuries",
  "Brucellosis"
)


outcome_lookup <- tibble(

  outcome_name =
    outcome_cols,

  outcome_category =
    c(

      "Malaria",
      "Tuberculosis and mycobacterial diseases",
      "HIV/AIDS",
      "Population and prevention",
      "Population and demographic outcomes",
      "Population and mortality outcomes",
      "Food-/water-borne infectious diseases",
      "Maternal and child health",
      "Nutrition and anaemia",
      "Respiratory infectious diseases (non-emerging)",
      "Other infectious diseases (non-emerging)",
      "Food-/water-borne infectious diseases",
      "Nutrition and anaemia",
      "Respiratory infectious diseases (non-emerging)",
      "Drug resistance / AMR",
      "Non-communicable diseases",
      "Non-communicable diseases",
      "Non-communicable diseases",
      "Non-communicable diseases",
      "Non-communicable diseases",
      "Non-communicable diseases",
      "Non-communicable diseases",
      "Respiratory infectious diseases (non-emerging)",
      "Respiratory infectious diseases (non-emerging)",
      "Syndromic / undifferentiated infectious illnesses",
      "Parasitic / helminthic infectious diseases",
      "Parasitic / helminthic infectious diseases",
      "Zoonotic infectious diseases (non-emerging, non-vector-borne)",
      "Respiratory infectious diseases (non-emerging)",
      "Emerging / epidemic-prone infectious diseases",
      "Respiratory infectious diseases (non-emerging)",
      "Respiratory infectious diseases (non-emerging)",
      "Emerging / epidemic-prone infectious diseases",
      "Zoonotic infectious diseases (non-emerging, non-vector-borne)",
      "Sensory / oral / other health outcomes",
      "Zoonotic infectious diseases (emerging/epidemic-prone)",
      "Zoonotic infectious diseases (emerging/epidemic-prone)",
      "Zoonotic infectious diseases (non-emerging, non-vector-borne)",
      "Non-malarial vector-borne infectious diseases",
      "Zoonotic infectious diseases (non-emerging, non-vector-borne)",
      "Food-/water-borne infectious diseases",
      "Food-/water-borne infectious diseases",
      "Food-/water-borne infectious diseases",
      "Food-/water-borne infectious diseases",
      "Zoonotic infectious diseases (emerging/epidemic-prone)",
      "Non-malarial vector-borne infectious diseases",
      "Non-malarial vector-borne infectious diseases",
      "Non-malarial vector-borne infectious diseases",
      "Non-malarial vector-borne infectious diseases",
      "Non-malarial vector-borne infectious diseases",
      "Non-malarial vector-borne infectious diseases",
      "Non-malarial vector-borne infectious diseases",
      "Non-malarial vector-borne infectious diseases",
      "Non-malarial vector-borne infectious diseases",
      "Non-malarial vector-borne infectious diseases",
      "Non-malarial vector-borne infectious diseases",
      "Respiratory infectious diseases (non-emerging)",
      "Parasitic / helminthic infectious diseases",
      "Syndromic / undifferentiated infectious illnesses",
      "Parasitic / helminthic infectious diseases",
      "Non-communicable diseases",
      "Other infectious diseases (non-emerging)",
      "Non-communicable diseases",
      "Clinical / examination / biomarkers",
      "Sensory / oral / other health outcomes",
      "Other infectious diseases (non-emerging)",
      "Zoonotic infectious diseases (non-emerging, non-vector-borne)",
      "Injuries / trauma",
      "Zoonotic infectious diseases (non-emerging, non-vector-borne)"
    )
)


infectious_categories <- c(

  "Malaria",
  "Tuberculosis and mycobacterial diseases",
  "HIV/AIDS",
  "Respiratory infectious diseases (non-emerging)",
  "Zoonotic infectious diseases (non-emerging, non-vector-borne)",
  "Zoonotic infectious diseases (emerging/epidemic-prone)",
  "Food-/water-borne infectious diseases",
  "Parasitic / helminthic infectious diseases",
  "Non-malarial vector-borne infectious diseases",
  "Emerging / epidemic-prone infectious diseases",
  "Syndromic / undifferentiated infectious illnesses",
  "Other infectious diseases (non-emerging)",
  "Drug resistance / AMR"
)
```

# 12. Clean cohort characteristics

```{r cohort-characteristics}
cohort_clean <-
  cohort_master %>%

  mutate(

    cohort_name =
      clean_text(
        cohort_name
      ),

    country_raw =
      clean_text(
        country
      ),

    country_clean =
      clean_text(
        country
      ),

    country_type =
      clean_country_type(
        country_clean
      ),

    region =
      classify_region(
        country_clean
      ),

    participant_size_raw =
      clean_text(
        participant_size
      ),

    participant_size_n =
      map_dbl(
        participant_size_raw,
        extract_sample_size_numeric
      ),

    participant_size_cat =
      categorise_sample_size(
        participant_size_n
      ),

    setting_raw =
      clean_text(
        setting
      ),

    setting_clean =
      clean_setting(
        setting
      ),

    start_year_num =
      suppressWarnings(
        as.numeric(
          clean_text(
            start_year
          )
        )
      ),

    start_decade =
      categorise_decade(
        start_year_num
      ),

    end_year_num =
      suppressWarnings(
        as.numeric(
          clean_text(
            end_year
          )
        )
      ),

    cohort_status =
      categorise_status(
        end_year_num
      ),

    hdss_clean =
      case_when(

        is_yes(hdss) ~
          "HDSS",

        !is.na(hdss) &
          clean_text(hdss) != "" ~
          "Non-HDSS",

        TRUE ~
          "Unclear"
      ),

    one_health_clean =
      case_when(

        is_yes(
          mention_one_health
        ) ~
          "Mentions One Health",

        !is.na(
          mention_one_health
        ) &

          clean_text(
            mention_one_health
          ) != "" ~

          "Does not mention One Health",

        TRUE ~
          "Unclear"
      ),

    data_sharing_clean =
      case_when(

        str_detect(
          tolower(
            clean_text(
              data_sharing_potential
            )
          ),
          paste0(
            "yes|available|open|request|",
            "upon request|application"
          )
        ) ~
          "Potentially shareable",

        str_detect(
          tolower(
            clean_text(
              data_sharing_potential
            )
          ),
          "no|restricted|not available|not shared"
        ) ~
          "Restricted/not shareable",

        TRUE ~
          "Unclear"
      ),

    age_category_clean2 =
      case_when(

        is_yes(
          age_all_ages
        ) ~
          "All ages",

        is_yes(
          age_children
        ) &
          is_yes(
            age_adults
          ) ~
          "All ages",

        is_yes(
          age_adolescents
        ) &
          !is_yes(
            age_adults
          ) &
          !is_yes(
            age_children
          ) ~
          "Children only",

        is_yes(
          age_children
        ) &
          !is_yes(
            age_adults
          ) ~
          "Children only",

        !is_yes(
          age_children
        ) &
          !is_yes(
            age_adolescents
          ) &
          is_yes(
            age_adults
          ) ~
          "Adults only",

        TRUE ~
          NA_character_
      ),

    sex_gender_clean2 =
      case_when(

        is_yes(
          sex_both
        ) ~
          "Both sexes",

        is_yes(
          sex_female
        ) &
          !is_yes(
            sex_male
          ) ~
          "Female only",

        is_yes(
          sex_male
        ) &
          !is_yes(
            sex_female
          ) ~
          "Male only",

        is_yes(
          sex_male
        ) &
          is_yes(
            sex_female
          ) ~
          "Both sexes",

        TRUE ~
          NA_character_
      )
  )


cat(
  "Rows in cohort_clean:",
  nrow(cohort_clean),
  "\n"
)


cat(
  "Unique cohort_id:",
  n_distinct(
    cohort_clean$cohort_id
  ),
  "\n"
)


cat(
  "Unique cohort_name:",
  n_distinct(
    cohort_clean$cohort_name
  ),
  "\n"
)


cohort_clean %>%

  count(
    cohort_name,
    sort = TRUE
  ) %>%

  filter(
    n > 1
  )
```

# 13. Long outcome dataset and follow-up

```{r outcome-long-followup}
outcome_cols_present <-
  intersect(
    outcome_cols,
    names(cohort_clean)
  )


outcome_long_clean <-
  cohort_clean %>%

  select(
    cohort_id,
    cohort_name,
    all_of(
      outcome_cols_present
    )
  ) %>%

  pivot_longer(

    cols =
      all_of(
        outcome_cols_present
      ),

    names_to =
      "outcome_name",

    values_to =
      "outcome_value_raw"
  ) %>%

  mutate(

    outcome_present =
      is_yes(
        outcome_value_raw
      )
  ) %>%

  filter(
    outcome_present
  ) %>%

  left_join(
    outcome_lookup,
    by = "outcome_name"
  ) %>%

  mutate(

    outcome_category =
      if_else(
        is.na(
          outcome_category
        ),
        "Other/unclear",
        outcome_category
      ),

    broad_outcome_domain =
      case_when(

        outcome_category %in%
          infectious_categories ~
          "Infectious disease",

        outcome_category ==
          "Non-communicable diseases" ~
          "Non-communicable diseases",

        outcome_category %in%
          c(
            "Population and prevention",
            "Population and demographic outcomes",
            "Population and mortality outcomes",
            "Maternal and child health",
            "Nutrition and anaemia"
          ) ~
          "Population / prevention / nutrition",

        outcome_category ==
          "Clinical / examination / biomarkers" ~
          "Clinical / examination",

        outcome_category %in%
          c(
            "Sensory / oral / other health outcomes",
            "Injuries / trauma"
          ) ~
          "Other health outcomes",

        TRUE ~
          "Other/unclear"
      )
  )


outcome_by_cohort <-
  outcome_long_clean %>%

  distinct(
    cohort_id,
    outcome_name,
    outcome_category
  ) %>%

  group_by(
    cohort_id
  ) %>%

  summarise(

    outcome_categories_clean =
      paste(
        sort(
          unique(
            outcome_category
          )
        ),
        collapse = "; "
      ),

    n_outcome_categories =
      n_distinct(
        outcome_category
      ),

    individual_outcomes_clean =
      paste(
        sort(
          unique(
            outcome_name
          )
        ),
        collapse = "; "
      ),

    n_individual_outcomes =
      n_distinct(
        outcome_name
      ),

    .groups =
      "drop"
  )


outcome_domain_by_cohort <-
  outcome_long_clean %>%

  distinct(
    cohort_id,
    broad_outcome_domain
  ) %>%

  group_by(
    cohort_id
  ) %>%

  summarise(

    broad_outcome_domains =
      paste(
        sort(
          unique(
            broad_outcome_domain
          )
        ),
        collapse = "; "
      ),

    .groups =
      "drop"
  )


outcome_details_clean <-
  outcome_details %>%

  mutate(

    cohort_name =
      clean_text(
        cohort_name
      ),

    Outcome =
      clean_text(
        Outcome
      ),

    followup_raw =
      `Follow Up  (Months -for specific outcome)`,

    followup_raw_clean =
      clean_text(
        followup_raw
      ),

    followup_months =
      map_dbl(
        followup_raw,
        parse_followup_months
      )
  )


# Follow-up information was extracted at the
# harmonised cohort-name level. Jointly reported
# cohort sites therefore receive the same
# source-level follow-up information.

followup_by_name <-
  outcome_details_clean %>%

  group_by(
    cohort_name
  ) %>%

  summarise(

    n_outcome_followup_records =
      sum(
        !is.na(
          followup_months
        )
      ),

    followup_min_months =
      ifelse(

        n_outcome_followup_records >
          0,

        min(
          followup_months,
          na.rm = TRUE
        ),

        NA_real_
      ),

    followup_max_months =
      ifelse(

        n_outcome_followup_records >
          0,

        max(
          followup_months,
          na.rm = TRUE
        ),

        NA_real_
      ),

    n_distinct_followup_values =
      n_distinct(
        followup_months[
          !is.na(
            followup_months
          )
        ]
      ),

    .groups =
      "drop"
  ) %>%

  mutate(

    followup_range_cat =
      categorise_followup_range(
        followup_min_months,
        followup_max_months
      ),

    followup_midpoint_months =
      ifelse(

        !is.na(
          followup_min_months
        ) &
          !is.na(
            followup_max_months
          ),

        (
          followup_min_months +
            followup_max_months
        ) / 2,

        NA_real_
      )
  )


cohort_characteristics <-
  cohort_clean %>%

  left_join(
    outcome_by_cohort,
    by = "cohort_id"
  ) %>%

  left_join(
    outcome_domain_by_cohort,
    by = "cohort_id"
  ) %>%

  left_join(
    followup_by_name,
    by = "cohort_name"
  )
```

# 14. Descriptive cohort-characteristics tables

```{r descriptive-tables}
table_region <-
  summ_cat(
    cohort_characteristics,
    region
  ) %>%

  mutate(
    variable =
      "Region"
  )


table_country_type <-
  summ_cat(
    cohort_characteristics,
    country_type
  ) %>%

  mutate(
    variable =
      "Country classification"
  )


table_sample_size <-
  summ_cat(
    cohort_characteristics,
    participant_size_cat
  ) %>%

  mutate(
    variable =
      "Participant size"
  )


table_start_decade <-
  summ_cat(
    cohort_characteristics,
    start_decade
  ) %>%

  mutate(
    variable =
      "Start decade"
  )


table_status <-
  summ_cat(
    cohort_characteristics,
    cohort_status
  ) %>%

  mutate(
    variable =
      "Cohort status"
  )


table_setting <-
  summ_cat(
    cohort_characteristics,
    setting_clean
  ) %>%

  mutate(
    variable =
      "Setting"
  )


table_data_sharing <-
  summ_cat(
    cohort_characteristics,
    data_sharing_clean
  ) %>%

  mutate(
    variable =
      "Data sharing potential"
  )


table_hdss <-
  summ_cat(
    cohort_characteristics,
    hdss_clean
  ) %>%

  mutate(
    variable =
      "HDSS status"
  )


table_one_health <-
  summ_cat(
    cohort_characteristics,
    one_health_clean
  ) %>%

  mutate(
    variable =
      "One Health mention"
  )


table_age <-
  summ_cat(
    cohort_characteristics,
    age_category_clean2
  ) %>%

  mutate(
    variable =
      "Age category"
  )


table_sex <-
  summ_cat(
    cohort_characteristics,
    sex_gender_clean2
  ) %>%

  mutate(
    variable =
      "Sex/gender category"
  )


table_interface <-
  summ_cat(
    cohort_characteristics,
    broad_interface_cleaned
  ) %>%

  mutate(
    variable =
      "Broad One Health interface"
  )


table_followup <-
  summ_cat(
    cohort_characteristics,
    followup_range_cat
  ) %>%

  mutate(
    variable =
      "Follow-up range"
  )


cohort_characteristics_table <-
  bind_rows(

    table_region,
    table_country_type,
    table_sample_size,
    table_start_decade,
    table_status,
    table_setting,
    table_data_sharing,
    table_hdss,
    table_one_health,
    table_age,
    table_sex,
    table_interface,
    table_followup
  ) %>%

  mutate(

    category =
      as.character(
        category
      ),

    `n (%)` =
      paste0(
        n,
        " (",
        percent,
        ")"
      )
  ) %>%

  select(
    variable,
    category,
    `n (%)`
  )


cohort_characteristics_table_ft <-
  cohort_characteristics_table %>%

  group_by(
    variable
  ) %>%

  mutate(

    variable_display =
      if_else(
        row_number() == 1,
        variable,
        ""
      )
  ) %>%

  ungroup() %>%

  select(

    Characteristic =
      variable_display,

    Category =
      category,

    `No. of cohorts (%)` =
      `n (%)`
  )


ft_cohort_characteristics <-
  flextable(
    cohort_characteristics_table_ft
  ) %>%

  theme_booktabs() %>%

  autofit() %>%

  bold(
    part = "header"
  ) %>%

  align(
    j = 1:2,
    align = "left",
    part = "all"
  ) %>%

  align(
    j = 3,
    align = "center",
    part = "all"
  ) %>%

  fontsize(
    size = 10,
    part = "all"
  ) %>%

  set_caption(
    "Table. Summary characteristics of included cohorts"
  )


ft_cohort_characteristics


method_summary <- tibble(

  Method =
    c(
      "Active surveillance",
      "Passive/sentinel surveillance",
      "Questionnaire/survey",
      "Clinical assessment",
      "Laboratory assessment",
      "Routine/registry data"
    ),

  n =
    c(

      sum(
        cohort_clean$active_surveillance_detection ==
          "Yes",
        na.rm = TRUE
      ),

      sum(
        cohort_clean$case_based_passive_sentinel_detection ==
          "Yes",
        na.rm = TRUE
      ),

      sum(
        cohort_clean$questionnaire_data ==
          "Yes",
        na.rm = TRUE
      ),

      sum(
        cohort_clean$clinical_data ==
          "Yes",
        na.rm = TRUE
      ),

      sum(
        cohort_clean$laboratory_data ==
          "Yes",
        na.rm = TRUE
      ),

      sum(
        cohort_clean$routine_data ==
          "Yes",
        na.rm = TRUE
      )
    )
) %>%

  mutate(

    percent =
      round(
        100 *
          n /
          nrow(cohort_clean),
        1
      )
  )


method_summary
```

# 15. Broad One Health interface figure

```{r interface-figure, fig.width=8, fig.height=4.5}
interface_summary <-
  cohort_clean %>%

  filter(
    broad_interface_cleaned %in%
      c(
        "Human-Environmental",
        "Human-Animal-Environmental"
      )
  ) %>%

  count(
    broad_interface_cleaned,
    name =
      "n_cohorts"
  ) %>%

  mutate(

    proportion =
      n_cohorts /
        nrow(cohort_clean),

    label =
      percent(
        proportion,
        accuracy = 0.1
      )
  )


p_interface <-
  ggplot(

    interface_summary,

    aes(

      x =
        proportion,

      y =
        fct_reorder(
          broad_interface_cleaned,
          proportion
        )
    )
  ) +

  geom_col(
    fill = "#08306b",
    width = 0.7
  ) +

  geom_text(

    aes(
      label =
        label
    ),

    hjust = -0.08,
    size = 4.5
  ) +

  scale_x_continuous(

    labels =
      percent_format(
        accuracy = 1
      ),

    limits =
      c(
        0,
        max(
          interface_summary$proportion
        ) * 1.15
      ),

    expand =
      expansion(
        mult = c(
          0,
          0.02
        )
      )
  ) +

  labs(
    x =
      "Proportion of cohorts",
    y =
      NULL
  ) +

  theme_minimal(
    base_size = 13
  ) +

  theme(
    panel.grid.major.y =
      element_blank(),
    panel.grid.minor =
      element_blank()
  )


p_interface
```

# 16. Individual outcomes, infectious outcomes, assays and human sample types

```{r outcome-assay-sample-summaries}
col_dark_blue <-
  "#004B7A"

col_mid_blue <-
  "#3B8BC2"

col_light_blue <-
  "#9CCBEA"

col_dark_red <-
  "#B30000"


n_total_cohorts <-
  n_distinct(
    cohort_clean$cohort_id
  )


outcomes_plot <-
  outcome_long_clean %>%

  mutate(

    outcome_name_plot =
      case_when(

        outcome_name ==
          "Pertusis" ~
          "Pertussis",

        outcome_name ==
          "Menengitis" ~
          "Meningitis",

        outcome_name ==
          "soil-transmitted helminths." ~
          "Soil-transmitted helminths",

        TRUE ~
          outcome_name
      )
  )


individual_outcome_counts <-
  outcomes_plot %>%

  distinct(
    cohort_id,
    outcome_name_plot
  ) %>%

  count(
    outcome_name_plot,
    name =
      "n_cohorts"
  ) %>%

  mutate(

    proportion =
      n_cohorts /
        n_total_cohorts,

    percent =
      100 *
        proportion
  ) %>%

  arrange(
    desc(
      proportion
    )
  )


infectious_outcome_counts <-
  outcomes_plot %>%

  filter(
    outcome_category %in%
      infectious_categories
  ) %>%

  distinct(
    cohort_id,
    outcome_name_plot
  ) %>%

  count(
    outcome_name_plot,
    name =
      "n_cohorts"
  ) %>%

  mutate(

    proportion =
      n_cohorts /
        n_total_cohorts,

    percent =
      100 *
        proportion
  ) %>%

  arrange(
    desc(
      proportion
    )
  )


missing_from_red <-
  outcomes_plot %>%

  filter(
    outcome_category %in%
      infectious_categories
  ) %>%

  distinct(
    outcome_name_plot
  ) %>%

  anti_join(

    infectious_outcome_counts %>%
      select(
        outcome_name_plot
      ),

    by =
      "outcome_name_plot"
  )


cat(
  "Infectious outcomes missing from red graph:",
  nrow(missing_from_red),
  "\n"
)


assay_long <-
  cohort_clean %>%

  select(

    cohort_id,

    assay_serology,
    assay_pcr,
    assay_microscopy,
    assay_culture,
    assay_elisa,
    assay_sequencing,
    assay_hematology,
    assay_biomarkers,
    assay_anthropometry,
    assay_verbal_autopsy,
    assay_questionnaire,
    assay_rdt,
    assay_imaging,
    assay_spirometry
  ) %>%

  pivot_longer(

    cols =
      -cohort_id,

    names_to =
      "assay_var",

    values_to =
      "assessed"
  ) %>%

  filter(
    is_yes(
      assessed
    )
  ) %>%

  mutate(

    assay =
      recode(

        assay_var,

        assay_serology =
          "Serology",

        assay_pcr =
          "PCR",

        assay_microscopy =
          "Microscopy",

        assay_culture =
          "Culture",

        assay_elisa =
          "ELISA",

        assay_sequencing =
          "Sequencing",

        assay_hematology =
          "Hematology",

        assay_biomarkers =
          "Biomarkers",

        assay_anthropometry =
          "Anthropometry",

        assay_verbal_autopsy =
          "Verbal autopsy",

        assay_questionnaire =
          "Questionnaire",

        assay_rdt =
          "Rapid diagnostic test",

        assay_imaging =
          "Imaging",

        assay_spirometry =
          "Spirometry"
      )
  )


assay_counts <-
  assay_long %>%

  distinct(
    cohort_id,
    assay
  ) %>%

  count(
    assay,
    name =
      "n_cohorts"
  ) %>%

  mutate(

    proportion =
      n_cohorts /
        n_total_cohorts,

    percent =
      100 *
        proportion
  ) %>%

  arrange(
    desc(
      proportion
    )
  )


sample_long_human <-
  cohort_clean %>%

  select(

    cohort_id,

    sample_blood,
    sample_serum,
    sample_plasma,
    sample_dbs,
    sample_urine,
    sample_stool,
    sample_sputum,
    sample_swab,
    sample_tissue,
    sample_placenta,
    sample_saliva,
    sample_hair
  ) %>%

  pivot_longer(

    cols =
      -cohort_id,

    names_to =
      "sample_var",

    values_to =
      "assessed"
  ) %>%

  filter(
    is_yes(
      assessed
    )
  ) %>%

  mutate(

    sample_type =
      recode(

        sample_var,

        sample_blood =
          "Blood",

        sample_serum =
          "Serum",

        sample_plasma =
          "Plasma",

        sample_dbs =
          "DBS",

        sample_urine =
          "Urine",

        sample_stool =
          "Stool",

        sample_sputum =
          "Sputum",

        sample_swab =
          "Swab",

        sample_tissue =
          "Tissue",

        sample_placenta =
          "Placenta",

        sample_saliva =
          "Saliva",

        sample_hair =
          "Hair"
      )
  )


sample_counts_human <-
  sample_long_human %>%

  distinct(
    cohort_id,
    sample_type
  ) %>%

  count(
    sample_type,
    name =
      "n_cohorts"
  ) %>%

  mutate(

    proportion =
      n_cohorts /
        n_total_cohorts,

    percent =
      100 *
        proportion
  ) %>%

  arrange(
    desc(
      proportion
    )
  )
```

```{r outcome-figures, fig.width=11, fig.height=8}
p_individual_outcomes <-
  ggplot(

    individual_outcome_counts,

    aes(

      x =
        proportion,

      y =
        fct_reorder(
          outcome_name_plot,
          proportion
        )
    )
  ) +

  geom_col(
    fill =
      col_light_blue,
    width =
      0.72
  ) +

  geom_text(

    aes(
      label =
        percent_format(
          accuracy = 0.1
        )(
          proportion
        )
    ),

    hjust =
      -0.08,

    size =
      3.8
  ) +

  scale_x_continuous(

    labels =
      percent_format(
        accuracy = 1
      ),

    limits =
      c(
        0,
        max(
          individual_outcome_counts$proportion
        ) * 1.18
      ),

    expand =
      expansion(
        mult =
          c(
            0,
            0.02
          )
      )
  ) +

  labs(

    x =
      "Proportion of cohorts",

    y =
      NULL
  ) +

  theme_minimal(
    base_size = 12
  ) +

  theme(

    panel.grid.major.y =
      element_blank(),

    panel.grid.minor =
      element_blank()
  )


p_individual_outcomes
```

```{r infectious-outcome-figure, fig.width=12, fig.height=8}
p_infectious_outcomes <-
  ggplot(

    infectious_outcome_counts,

    aes(

      x =
        proportion,

      y =
        fct_reorder(
          outcome_name_plot,
          proportion
        )
    )
  ) +

  geom_col(

    fill =
      col_dark_red,

    width =
      0.72
  ) +

  geom_text(

    aes(
      label =
        percent_format(
          accuracy = 0.1
        )(
          proportion
        )
    ),

    hjust =
      -0.08,

    size =
      3.8
  ) +

  scale_x_continuous(

    labels =
      percent_format(
        accuracy = 1
      ),

    limits =
      c(
        0,
        max(
          infectious_outcome_counts$proportion
        ) * 1.18
      ),

    expand =
      expansion(
        mult =
          c(
            0,
            0.02
          )
      )
  ) +

  labs(

    x =
      "Proportion of cohorts",

    y =
      NULL
  ) +

  theme_minimal(
    base_size = 12
  ) +

  theme(

    panel.grid.major.y =
      element_blank(),

    panel.grid.minor =
      element_blank()
  )


p_infectious_outcomes
```

```{r assay-sample-figures, fig.width=9, fig.height=6}
p_assays <-
  ggplot(

    assay_counts,

    aes(

      x =
        proportion,

      y =
        fct_reorder(
          assay,
          proportion
        )
    )
  ) +

  geom_col(

    fill =
      col_light_blue,

    width =
      0.72
  ) +

  geom_text(

    aes(
      label =
        percent_format(
          accuracy = 0.1
        )(
          proportion
        )
    ),

    hjust =
      -0.08,

    size =
      4
  ) +

  scale_x_continuous(

    labels =
      percent_format(
        accuracy = 1
      ),

    limits =
      c(
        0,
        max(
          assay_counts$proportion
        ) * 1.18
      )
  ) +

  labs(
    x =
      "Proportion of cohorts",
    y =
      NULL
  ) +

  theme_minimal(
    base_size = 12
  ) +

  theme(
    panel.grid.major.y =
      element_blank()
  )


p_sample_types_human <-
  ggplot(

    sample_counts_human,

    aes(

      x =
        proportion,

      y =
        fct_reorder(
          sample_type,
          proportion
        )
    )
  ) +

  geom_col(

    fill =
      col_mid_blue,

    width =
      0.72
  ) +

  geom_text(

    aes(
      label =
        percent_format(
          accuracy = 0.1
        )(
          proportion
        )
    ),

    hjust =
      -0.08,

    size =
      4
  ) +

  scale_x_continuous(

    labels =
      percent_format(
        accuracy = 1
      ),

    limits =
      c(
        0,
        max(
          sample_counts_human$proportion
        ) * 1.18
      )
  ) +

  labs(
    x =
      "Proportion of cohorts",
    y =
      NULL
  ) +

  theme_minimal(
    base_size = 12
  ) +

  theme(
    panel.grid.major.y =
      element_blank()
  )


p_assays

p_sample_types_human
```

# 17. Data-integration matrix and score

```{r integration-matrix}
integration_scored <-
  cohort_clean %>%

  mutate(

    A_human_domain =
      human_outcome ==
        "Yes",

    B_animal_domain =
      animal_domain_cleaned ==
        "Yes",

    C_environment_domain =
      environmental_domain_cleaned ==
        "Yes",

    D_human_sampling =
      human_sampling_cleaned ==
        "Yes",

    E_animal_sampling =
      animal_sampling_cleaned ==
        "Yes",

    F_environment_sampling =
      environmental_sampling_cleaned ==
        "Yes" |

        environmental_monitoring_cleaned ==
          "Yes",

    G_climate_data =
      climate_data_cleaned ==
        "Yes",

    H_questionnaire =
      questionnaire_data ==
        "Yes",

    I_clinical =
      clinical_data ==
        "Yes",

    J_laboratory =
      laboratory_data ==
        "Yes",

    K_biobanking =
      !is.na(
        biobank_samples
      ) &

        !tolower(
          trimws(
            as.character(
              biobank_samples
            )
          )
        ) %in%
        c(
          "",
          "no",
          "n",
          "none",
          "none reported",
          "not reported",
          "not specified",
          "na",
          "n/a"
        ),

    L_registry =
      routine_data ==
        "Yes" |

        tolower(
          as.character(
            hdss
          )
        ) %in%
        c(
          "yes",
          "y",
          "true",
          "1"
        ),

    M_passive_detection =
      case_based_passive_sentinel_detection ==
        "Yes",

    N_verbal_autopsy =
      assay_verbal_autopsy ==
        "Yes" |

        special_deceased ==
          "Yes",

    O_active_surveillance =
      active_surveillance_detection ==
        "Yes",

    P_gis =
      `Geolocation (GPS)` ==
        "Yes"
  ) %>%

  mutate(

    data_integration_score =
      rowSums(

        cbind(

          A_human_domain,
          B_animal_domain,
          C_environment_domain,
          D_human_sampling,
          E_animal_sampling,
          F_environment_sampling,
          G_climate_data,
          H_questionnaire,
          I_clinical,
          J_laboratory,
          K_biobanking,
          L_registry,
          M_passive_detection,
          N_verbal_autopsy,
          O_active_surveillance,
          P_gis
        ),

        na.rm = TRUE
      )
  )


integration_data <-
  integration_scored %>%

  select(

    A_human_domain,
    B_animal_domain,
    C_environment_domain,
    D_human_sampling,
    E_animal_sampling,
    F_environment_sampling,
    G_climate_data,
    H_questionnaire,
    I_clinical,
    J_laboratory,
    K_biobanking,
    L_registry,
    M_passive_detection,
    N_verbal_autopsy,
    O_active_surveillance,
    P_gis
  )


integration_matrix <-
  matrix(

    0,

    nrow =
      ncol(
        integration_data
      ),

    ncol =
      ncol(
        integration_data
      ),

    dimnames =
      list(
        names(
          integration_data
        ),
        names(
          integration_data
        )
      )
  )


for (
  i in seq_len(
    ncol(
      integration_data
    )
  )
) {

  for (
    j in seq_len(
      ncol(
        integration_data
      )
    )
  ) {

    integration_matrix[
      i,
      j
    ] <-

      sum(

        integration_data[[i]] &
          integration_data[[j]],

        na.rm =
          TRUE
      )
  }
}


data_type_labels <- c(

  A_human_domain =
    "Human domain",

  B_animal_domain =
    "Animal domain",

  C_environment_domain =
    "Environmental domain",

  D_human_sampling =
    "Human biospecimen\nsampling",

  E_animal_sampling =
    "Animal sampling",

  F_environment_sampling =
    "Environmental sampling/\nmonitoring",

  G_climate_data =
    "Climate data",

  H_questionnaire =
    "Questionnaire/\nsurvey",

  I_clinical =
    "Clinical examination",

  J_laboratory =
    "Laboratory analysis",

  K_biobanking =
    "Biobanking",

  L_registry =
    "Registry/routine",

  M_passive_detection =
    "Passive surveillance",

  N_verbal_autopsy =
    "Verbal autopsy",

  O_active_surveillance =
    "Active surveillance",

  P_gis =
    "GIS/spatial data"
)


matrix_long <-
  as.data.frame(
    integration_matrix
  ) %>%

  rownames_to_column(
    "row_type"
  ) %>%

  pivot_longer(

    cols =
      -row_type,

    names_to =
      "col_type",

    values_to =
      "count"
  )


row_denominators <-
  diag(
    integration_matrix
  )


names(
  row_denominators
) <-
  rownames(
    integration_matrix
  )


matrix_long <-
  matrix_long %>%

  mutate(

    row_denom =
      row_denominators[
        row_type
      ],

    proportion =
      ifelse(
        row_denom > 0,
        count /
          row_denom,
        NA_real_
      ),

    proportion_label =
      case_when(

        is.na(
          proportion
        ) ~
          "",

        proportion ==
          1 ~
          "1",

        TRUE ~
          sprintf(
            "%.2f",
            proportion
          )
      ),

    row_type =
      factor(
        row_type,
        levels =
          rev(
            names(
              data_type_labels
            )
          )
      ),

    col_type =
      factor(
        col_type,
        levels =
          names(
            data_type_labels
          )
      )
  )


integration_type_score <-
  round(

    rowSums(
      integration_matrix,
      na.rm = TRUE
    ) /

      nrow(
        integration_data
      ),

    2
  )


integration_table <-
  as.data.frame(
    integration_matrix
  ) %>%

  rownames_to_column(
    "Data type"
  )


integration_table$`Data type` <-
  unname(
    data_type_labels[
      integration_table$`Data type`
    ]
  )


names(
  integration_table
)[
  2:ncol(
    integration_table
  )
] <-

  unname(

    data_type_labels[
      names(
        integration_table
      )[
        2:ncol(
          integration_table
        )
      ]
    ]
  )


integration_table$Score <-
  integration_type_score


ft_integration_matrix <-
  flextable(
    integration_table
  ) %>%

  theme_booktabs() %>%

  autofit() %>%

  bold(
    part =
      "header"
  ) %>%

  align(
    j = 1,
    align = "left",
    part = "all"
  ) %>%

  align(
    j =
      2:ncol(
        integration_table
      ),
    align =
      "center",
    part =
      "all"
  ) %>%

  fontsize(
    size = 9,
    part = "all"
  ) %>%

  bold(
    j = "Score",
    part = "body"
  ) %>%

  set_caption(
    "Table. Data-type co-occurrence and integration score"
  )


ft_integration_matrix
```

```{r integration-heatmap, fig.width=12, fig.height=9}
p_heat <-
  ggplot(

    matrix_long,

    aes(

      x =
        col_type,

      y =
        row_type,

      fill =
        proportion
    )
  ) +

  geom_tile(

    colour =
      "white",

    linewidth =
      0.45
  ) +

  geom_text(

    aes(

      label =
        proportion_label,

      colour =
        proportion >
          0.55
    ),

    size =
      3
  ) +

  scale_colour_manual(

    values =
      c(
        `FALSE` =
          "black",
        `TRUE` =
          "white"
      ),

    guide =
      "none"
  ) +

  scale_fill_gradient(

    low =
      "#EAF3FB",

    high =
      "#004B7A",

    limits =
      c(
        0,
        1
      ),

    labels =
      percent_format(
        accuracy = 1
      ),

    name =
      "Proportion"
  ) +

  scale_x_discrete(

    position =
      "top",

    labels =
      data_type_labels
  ) +

  scale_y_discrete(

    labels =
      data_type_labels
  ) +

  labs(
    x = NULL,
    y = NULL
  ) +

  theme_minimal(
    base_size =
      11
  ) +

  theme(

    panel.grid =
      element_blank(),

    axis.text.x.top =
      element_text(
        angle = 90,
        hjust = 0,
        vjust = 0.5,
        size = 8.5,
        colour = "black"
      ),

    axis.text.y =
      element_text(
        size = 8.5,
        colour = "black"
      ),

    axis.text.x.bottom =
      element_blank(),

    axis.ticks =
      element_blank()
  )


p_heat
```

# 18. Multivariable Poisson regression of data-integration score

Disease outcome categories are not included as explanatory variables. The
model includes region, setting, start decade, HDSS status and participant
size.

```{r regression-model}
analytic_df <-
  integration_scored %>%

  mutate(

    start_decade_collapsed =
      case_when(

        start_decade %in%
          c(
            "<=1970s",
            "1980s"
          ) ~
          "<=1980s",

        start_decade ==
          "1990s" ~
          "1990s",

        start_decade ==
          "2000s" ~
          "2000s",

        start_decade ==
          "2010s" ~
          "2010s",

        start_decade ==
          "2020s" ~
          "2020s",

        TRUE ~
          NA_character_
      ),

    region =
      factor(

        region,

        levels =
          c(
            "East Africa",
            "Central Africa",
            "Multinational",
            "North Africa",
            "Southern Africa",
            "West Africa"
          )
      ),

    setting_clean =
      factor(

        setting_clean,

        levels =
          c(
            "Urban",
            "Rural",
            "Mixed rural and urban"
          )
      ),

    start_decade_collapsed =
      factor(

        start_decade_collapsed,

        levels =
          c(
            "<=1980s",
            "1990s",
            "2000s",
            "2010s",
            "2020s"
          )
      ),

    hdss_binary =
      case_when(

        hdss_clean ==
          "HDSS" ~
          "HDSS",

        hdss_clean ==
          "Non-HDSS" ~
          "Non-HDSS",

        TRUE ~
          NA_character_
      ),

    hdss_binary =
      factor(

        hdss_binary,

        levels =
          c(
            "Non-HDSS",
            "HDSS"
          )
      ),

    participant_size_cat =
      factor(

        participant_size_cat,

        levels =
          c(
            "<1,000",
            "1,000-19,999",
            "20,000-99,999",
            ">=100,000"
          )
      )
  )


model_df <-
  analytic_df %>%

  filter(

    !is.na(
      data_integration_score
    ),

    !is.na(
      region
    ),

    !is.na(
      setting_clean
    ),

    !is.na(
      start_decade_collapsed
    ),

    !is.na(
      hdss_binary
    ),

    !is.na(
      participant_size_cat
    )
  )


cat(
  "Regression model N =",
  nrow(model_df),
  "\n"
)


m_pois <-
  glm(

    data_integration_score ~

      region +

      setting_clean +

      start_decade_collapsed +

      hdss_binary +

      participant_size_cat,

    data =
      model_df,

    family =
      poisson(
        link = "log"
      )
  )


vcov_robust <-
  sandwich::vcovHC(
    m_pois,
    type = "HC0"
  )


robust_test <-
  lmtest::coeftest(
    m_pois,
    vcov. =
      vcov_robust
  )


model_results <-
  tibble(

    term =
      rownames(
        robust_test
      ),

    estimate =
      robust_test[
        ,
        "Estimate"
      ],

    std.error =
      robust_test[
        ,
        "Std. Error"
      ],

    statistic =
      robust_test[
        ,
        "z value"
      ],

    p.value =
      robust_test[
        ,
        "Pr(>|z|)"
      ]
  ) %>%

  mutate(

    rr =
      exp(
        estimate
      ),

    conf.low =
      exp(
        estimate -
          1.96 *
          std.error
      ),

    conf.high =
      exp(
        estimate +
          1.96 *
          std.error
      )
  )


model_results_clean <-
  model_results %>%

  filter(
    term !=
      "(Intercept)"
  ) %>%

  mutate(

    variable =
      case_when(

        str_detect(
          term,
          "^region"
        ) ~
          "Region",

        str_detect(
          term,
          "^setting_clean"
        ) ~
          "Setting",

        str_detect(
          term,
          "^start_decade_collapsed"
        ) ~
          "Start decade",

        str_detect(
          term,
          "^hdss_binary"
        ) ~
          "HDSS status",

        str_detect(
          term,
          "^participant_size_cat"
        ) ~
          "Participant size",

        TRUE ~
          NA_character_
      ),

    level =
      case_when(

        str_detect(
          term,
          "^region"
        ) ~
          str_remove(
            term,
            "^region"
          ),

        str_detect(
          term,
          "^setting_clean"
        ) ~
          str_remove(
            term,
            "^setting_clean"
          ),

        str_detect(
          term,
          "^start_decade_collapsed"
        ) ~
          str_remove(
            term,
            "^start_decade_collapsed"
          ),

        str_detect(
          term,
          "^hdss_binary"
        ) ~
          str_remove(
            term,
            "^hdss_binary"
          ),

        str_detect(
          term,
          "^participant_size_cat"
        ) ~
          str_remove(
            term,
            "^participant_size_cat"
          ),

        TRUE ~
          NA_character_
      ),

    p_value_label =
      ifelse(

        p.value <
          0.001,

        "<0.001",

        sprintf(
          "%.3f",
          p.value
        )
      )
  )


model_results_clean %>%

  select(
    variable,
    level,
    rr,
    conf.low,
    conf.high,
    p.value
  )
```

```{r regression-table}
count_region <-
  model_df %>%

  count(
    region,
    name = "n"
  ) %>%

  transmute(

    variable =
      "Region",

    level =
      as.character(
        region
      ),

    n_label =
      paste0(
        "(n=",
        n,
        ")"
      )
  )


count_setting <-
  model_df %>%

  count(
    setting_clean,
    name = "n"
  ) %>%

  transmute(

    variable =
      "Setting",

    level =
      as.character(
        setting_clean
      ),

    n_label =
      paste0(
        "(n=",
        n,
        ")"
      )
  )


count_decade <-
  model_df %>%

  count(
    start_decade_collapsed,
    name = "n"
  ) %>%

  transmute(

    variable =
      "Start decade",

    level =
      as.character(
        start_decade_collapsed
      ),

    n_label =
      paste0(
        "(n=",
        n,
        ")"
      )
  )


count_hdss <-
  model_df %>%

  count(
    hdss_binary,
    name = "n"
  ) %>%

  transmute(

    variable =
      "HDSS status",

    level =
      as.character(
        hdss_binary
      ),

    n_label =
      paste0(
        "(n=",
        n,
        ")"
      )
  )


count_size <-
  model_df %>%

  count(
    participant_size_cat,
    name = "n"
  ) %>%

  transmute(

    variable =
      "Participant size",

    level =
      as.character(
        participant_size_cat
      ),

    n_label =
      paste0(
        "(n=",
        n,
        ")"
      )
  )


counts_all <-
  bind_rows(
    count_region,
    count_setting,
    count_decade,
    count_hdss,
    count_size
  )


reference_levels <-
  tibble(

    variable =
      c(
        "Region",
        "Setting",
        "Start decade",
        "HDSS status",
        "Participant size"
      ),

    level =
      c(
        "East Africa",
        "Urban",
        "<=1980s",
        "Non-HDSS",
        "<1,000"
      )
  )


estimates_table <-
  model_results_clean %>%

  transmute(

    variable,

    level,

    RR =
      sprintf(
        "%.2f",
        rr
      ),

    `95% CI` =
      sprintf(
        "%.2f, %.2f",
        conf.low,
        conf.high
      ),

    `P-value` =
      p_value_label
  )


table_body <-
  counts_all %>%

  left_join(

    estimates_table,

    by =
      c(
        "variable",
        "level"
      )
  ) %>%

  left_join(

    reference_levels %>%
      mutate(
        reference =
          TRUE
      ),

    by =
      c(
        "variable",
        "level"
      )
  ) %>%

  mutate(

    RR =
      if_else(
        reference %in%
          TRUE,
        "Reference",
        RR
      ),

    `95% CI` =
      if_else(
        reference %in%
          TRUE,
        "",
        `95% CI`
      ),

    `P-value` =
      if_else(
        reference %in%
          TRUE,
        "",
        `P-value`
      )
  )


make_section <- function(
    data,
    section
) {

  bind_rows(

    tibble(

      variable =
        section,

      level =
        "",

      n_label =
        "",

      RR =
        "",

      `95% CI` =
        "",

      `P-value` =
        ""
    ),

    data %>%

      filter(
        variable ==
          section
      ) %>%

      select(
        variable,
        level,
        n_label,
        RR,
        `95% CI`,
        `P-value`
      )
  )
}


table_display <-
  bind_rows(

    make_section(
      table_body,
      "Region"
    ),

    make_section(
      table_body,
      "Setting"
    ),

    make_section(
      table_body,
      "Start decade"
    ),

    make_section(
      table_body,
      "HDSS status"
    ),

    make_section(
      table_body,
      "Participant size"
    )
  ) %>%

  mutate(

    Characteristic =
      if_else(
        level == "",
        variable,
        level
      ),

    `(n)` =
      n_label
  ) %>%

  select(
    Characteristic,
    `(n)`,
    RR,
    `95% CI`,
    `P-value`
  )


ft_regression_style <-
  flextable(
    table_display
  ) %>%

  theme_booktabs() %>%

  autofit() %>%

  fontsize(
    size = 10,
    part = "all"
  ) %>%

  bold(
    part =
      "header"
  ) %>%

  align(
    j = 1,
    align = "left",
    part = "all"
  ) %>%

  align(
    j = 2:5,
    align = "center",
    part = "all"
  ) %>%

  bg(

    i =
      ~ `(n)` == "" &
      RR == "" &
      `95% CI` == "" &
      `P-value` == "",

    bg =
      "#D9D9D9",

    part =
      "body"
  ) %>%

  bold(

    i =
      ~ `(n)` == "" &
      RR == "" &
      `95% CI` == "" &
      `P-value` == "",

    bold =
      TRUE,

    part =
      "body"
  ) %>%

  set_caption(

    paste0(
      "Table. Multivariable Poisson regression of factors associated with ",
      "data integration score (N=",
      nrow(model_df),
      ")."
    )
  )


ft_regression_style
```

```{r regression-forest, fig.width=10, fig.height=7}
level_order_tbl <-
  tibble::tribble(

    ~variable,
    ~level,
    ~block_order,
    ~order_within,

    "Region",
    "East Africa",
    1,
    1,

    "Region",
    "Central Africa",
    1,
    2,

    "Region",
    "Multinational",
    1,
    3,

    "Region",
    "North Africa",
    1,
    4,

    "Region",
    "Southern Africa",
    1,
    5,

    "Region",
    "West Africa",
    1,
    6,

    "Setting",
    "Urban",
    2,
    1,

    "Setting",
    "Rural",
    2,
    2,

    "Setting",
    "Mixed rural and urban",
    2,
    3,

    "Start decade",
    "<=1980s",
    3,
    1,

    "Start decade",
    "1990s",
    3,
    2,

    "Start decade",
    "2000s",
    3,
    3,

    "Start decade",
    "2010s",
    3,
    4,

    "Start decade",
    "2020s",
    3,
    5,

    "HDSS status",
    "Non-HDSS",
    4,
    1,

    "HDSS status",
    "HDSS",
    4,
    2,

    "Participant size",
    "<1,000",
    5,
    1,

    "Participant size",
    "1,000-19,999",
    5,
    2,

    "Participant size",
    "20,000-99,999",
    5,
    3,

    "Participant size",
    ">=100,000",
    5,
    4
  )


forest_ref_rows <-
  reference_levels %>%

  transmute(

    variable,

    level,

    label_short =
      paste0(
        level,
        " (Ref)"
      ),

    rr =
      1,

    conf.low =
      1,

    conf.high =
      1,

    p.value =
      NA_real_,

    p_label =
      "Ref",

    sig =
      FALSE,

    is_reference =
      TRUE
  )


forest_est_rows <-
  model_results_clean %>%

  mutate(

    label_short =
      paste0(
        "   ",
        level
      ),

    p_label =
      case_when(

        p.value <
          0.001 ~
          "<0.001*",

        p.value <
          0.05 ~
          paste0(
            sprintf(
              "%.3f",
              p.value
            ),
            "*"
          ),

        TRUE ~
          sprintf(
            "%.3f",
            p.value
          )
      ),

    sig =
      p.value <
        0.05,

    is_reference =
      FALSE
  ) %>%

  select(

    variable,
    level,
    label_short,
    rr,
    conf.low,
    conf.high,
    p.value,
    p_label,
    sig,
    is_reference
  )


forest_df <-
  bind_rows(
    forest_ref_rows,
    forest_est_rows
  ) %>%

  left_join(

    level_order_tbl,

    by =
      c(
        "variable",
        "level"
      )
  ) %>%

  arrange(
    block_order,
    order_within
  ) %>%

  mutate(
    y_pos =
      rev(
        seq_len(
          n()
        )
      )
  )


xmax_plot <-
  max(

    1.8,

    max(
      forest_df$conf.high,
      na.rm = TRUE
    ) * 1.35
  )


x_pval <-
  xmax_plot *
  0.98


p_forest <-
  ggplot(

    forest_df,

    aes(
      x = rr,
      y = y_pos
    )
  ) +

  geom_vline(

    xintercept =
      1,

    linetype =
      "dotted",

    colour =
      "red",

    linewidth =
      0.8
  ) +

  geom_errorbar(

    data =
      forest_df %>%
      filter(
        !is_reference
      ),

    aes(
      xmin =
        conf.low,
      xmax =
        conf.high
    ),

    width =
      0.18,

    colour =
      "grey45",

    linewidth =
      0.6
  ) +

  geom_point(

    aes(
      fill =
        sig,
      shape =
        is_reference
    ),

    size =
      2.8,

    colour =
      "#004B7A",

    stroke =
      0.4
  ) +

  geom_text(

    aes(
      x =
        x_pval,
      label =
        p_label
    ),

    hjust =
      1,

    size =
      3.2
  ) +

  scale_shape_manual(

    values =
      c(
        `TRUE` =
          22,
        `FALSE` =
          21
      ),

    guide =
      "none"
  ) +

  scale_fill_manual(

    values =
      c(
        `TRUE` =
          "#B30000",
        `FALSE` =
          "#004B7A"
      ),

    guide =
      "none"
  ) +

  scale_y_continuous(

    breaks =
      forest_df$y_pos,

    labels =
      forest_df$label_short,

    expand =
      expansion(
        mult =
          c(
            0.02,
            0.06
          )
      )
  ) +

  scale_x_continuous(

    limits =
      c(
        0,
        xmax_plot
      ),

    breaks =
      pretty(
        c(
          0,
          xmax_plot
        ),
        n = 6
      )
  ) +

  labs(

    x =
      "Rate Ratio (RR)",

    y =
      NULL
  ) +

  theme_minimal(
    base_size =
      12
  ) +

  theme(

    panel.grid.major.y =
      element_blank(),

    panel.grid.minor =
      element_blank(),

    axis.text.y =
      element_text(
        size = 9
      ),

    plot.margin =
      margin(
        10,
        35,
        10,
        10
      )
  ) +

  coord_cartesian(
    clip =
      "off"
  )


p_forest
```

# 19. Temporal trends

```{r time-trends, fig.width=9, fig.height=5.5}
time_data <-
  cohort_clean %>%

  filter(
    !is.na(
      start_year_num
    )
  ) %>%

  mutate(

    decade =
      floor(
        start_year_num /
          10
      ) * 10,

    interface_clean =
      broad_interface_cleaned
  )


time_all <-
  time_data %>%

  count(
    decade,
    name = "n"
  ) %>%

  mutate(
    series =
      "All cohorts"
  )


time_he <-
  time_data %>%

  filter(
    interface_clean ==
      "Human-Environmental"
  ) %>%

  count(
    decade,
    name = "n"
  ) %>%

  mutate(
    series =
      "Human-Environmental"
  )


time_hae <-
  time_data %>%

  filter(
    interface_clean ==
      "Human-Animal-Environmental"
  ) %>%

  count(
    decade,
    name = "n"
  ) %>%

  mutate(
    series =
      "Human-Animal-Environmental"
  )


time_line_data <-
  bind_rows(
    time_all,
    time_he,
    time_hae
  ) %>%

  mutate(

    series =
      factor(

        series,

        levels =
          c(
            "All cohorts",
            "Human-Environmental",
            "Human-Animal-Environmental"
          )
      )
  )


line_palette <- c(

  "All cohorts" =
    "#004B7A",

  "Human-Environmental" =
    "#3B8BC2",

  "Human-Animal-Environmental" =
    "#B30000"
)


p_time_line <-
  ggplot(

    time_line_data,

    aes(

      x =
        decade,

      y =
        n,

      colour =
        series,

      group =
        series
    )
  ) +

  geom_line(
    linewidth =
      1.2
  ) +

  geom_point(
    size =
      3
  ) +

  scale_color_manual(
    values =
      line_palette
  ) +

  scale_x_continuous(

    breaks =
      seq(
        min(
          time_line_data$decade
        ),
        max(
          time_line_data$decade
        ),
        by = 10
      )
  ) +

  labs(

    x =
      "Start decade",

    y =
      "Number of cohorts",

    colour =
      NULL
  ) +

  theme_minimal(
    base_size =
      13
  ) +

  theme(
    panel.grid.minor =
      element_blank()
  )


p_time_line
```

# 20. Geographic distribution

Multinational cohorts are counted once in each African country in which
they were conducted. `cohort_id`, rather than harmonised cohort name, is
used so that distinct cohort/site records are preserved.

```{r geographic-data}
country_long <-
  cohort_clean %>%

  transmute(

    cohort_id,

    cohort_name,

    country =
      str_squish(
        country_raw
      ),

    broad_interface_cleaned
  ) %>%

  mutate(

    country =
      str_replace_all(
        country,
        "\u2019",
        "'"
      ),

    country =
      case_when(

        str_detect(
          country,
          regex(
            "African sites include",
            ignore_case = TRUE
          )
        ) ~

          str_replace(
            country,
            regex(
              "^.*African sites include\\s*",
              ignore_case = TRUE
            ),
            ""
          ),

        str_detect(
          country,
          regex(
            "^Multinational\\s*\\(",
            ignore_case = TRUE
          )
        ) ~

          str_replace(
            country,
            regex(
              "^Multinational\\s*\\(([^)]*)\\).*$",
              ignore_case = TRUE
            ),
            "\\1"
          ),

        TRUE ~
          country
      ),

    country =
      str_remove(
        country,
        "[\\.]+$"
      )
  ) %>%

  filter(
    !is.na(country),
    country != ""
  ) %>%

  separate_rows(

    country,

    sep =
      "\\s*[,;|/]\\s*|\\s+and\\s+"
  ) %>%

  mutate(

    country =
      str_squish(
        country
      ),

    country =
      case_when(

        country %in%
          c(
            "The Gambia",
            "Gambia"
          ) ~
          "Gambia",

        country %in%
          c(
            "Côte d'Ivoire",
            "Cote d'Ivoire",
            "Ivory Coast"
          ) ~
          "Côte d'Ivoire",

        country %in%
          c(
            "Democratic Republic of Congo",
            "Democratic Republic of the Congo",
            "DR Congo",
            "DRC"
          ) ~
          "Democratic Republic of the Congo",

        country %in%
          c(
            "Republic of Congo",
            "Congo Republic",
            "Congo"
          ) ~
          "Republic of the Congo",

        TRUE ~
          country
      ),

    iso3 =
      countrycode(

        country,

        origin =
          "country.name",

        destination =
          "iso3c"
      ),

    iso3 =
      case_when(

        country ==
          "Côte d'Ivoire" ~
          "CIV",

        country ==
          "Democratic Republic of the Congo" ~
          "COD",

        country ==
          "Republic of the Congo" ~
          "COG",

        country ==
          "Gambia" ~
          "GMB",

        TRUE ~
          iso3
      )
  ) %>%

  filter(
    !is.na(
      iso3
    )
  ) %>%

  distinct(
    cohort_id,
    iso3,
    .keep_all =
      TRUE
  )


coh_total <-
  country_long %>%

  distinct(
    cohort_id,
    iso3
  ) %>%

  count(
    iso3,
    name =
      "n_cohorts"
  ) %>%

  mutate(
    panel =
      "All cohorts\n(Human–Environmental data)"
  )


coh_hae <-
  country_long %>%

  filter(
    broad_interface_cleaned ==
      "Human-Animal-Environmental"
  ) %>%

  distinct(
    cohort_id,
    iso3
  ) %>%

  count(
    iso3,
    name =
      "n_cohorts"
  ) %>%

  mutate(
    panel =
      "Human–Animal–Environmental cohorts"
  )


panel_levels <- c(

  "All cohorts\n(Human–Environmental data)",

  "Human–Animal–Environmental cohorts"
)


coh_counts_facets <-
  bind_rows(
    coh_total,
    coh_hae
  ) %>%

  mutate(

    panel =
      factor(
        panel,
        levels =
          panel_levels
      )
  )


africa_sf <-
  rnaturalearth::ne_countries(

    scale =
      "medium",

    continent =
      "Africa",

    returnclass =
      "sf"
  ) %>%

  select(
    name_long,
    adm0_a3,
    geometry
  ) %>%

  rename(
    name =
      name_long,
    iso3 =
      adm0_a3
  )


grid_all <-
  tidyr::expand_grid(

    iso3 =
      africa_sf$iso3,

    panel =
      factor(
        panel_levels,
        levels =
          panel_levels
      )
  )


africa_facets <-
  africa_sf %>%

  st_drop_geometry() %>%

  select(
    iso3,
    name
  ) %>%

  distinct() %>%

  right_join(
    grid_all,
    by =
      "iso3"
  ) %>%

  left_join(

    coh_counts_facets,

    by =
      c(
        "iso3",
        "panel"
      )
  ) %>%

  mutate(

    n_cohorts =
      replace_na(
        n_cohorts,
        0L
      ),

    fill_value =
      ifelse(
        n_cohorts == 0,
        NA,
        n_cohorts
      )
  ) %>%

  left_join(

    africa_sf,

    by =
      c(
        "iso3",
        "name"
      )
  ) %>%

  st_as_sf()


label_points <-
  africa_sf %>%

  st_point_on_surface() %>%

  mutate(

    X =
      st_coordinates(.)[
        ,
        1
      ],

    Y =
      st_coordinates(.)[
        ,
        2
      ]
  ) %>%

  st_drop_geometry() %>%

  select(
    iso3,
    name,
    X,
    Y
  )


label_df <-
  africa_facets %>%

  st_drop_geometry() %>%

  filter(
    n_cohorts > 0
  ) %>%

  left_join(

    label_points,

    by =
      c(
        "iso3",
        "name"
      )
  ) %>%

  mutate(

    short_name =
      recode(

        name,

        "Côte d'Ivoire" =
          "Côte d’Ivoire",

        "Guinea-Bissau" =
          "G.-Bissau",

        "Central African Republic" =
          "C.A.R.",

        "Democratic Republic of the Congo" =
          "DR Congo",

        "Republic of the Congo" =
          "Congo",

        "Equatorial Guinea" =
          "Eq. Guinea",

        .default =
          name
      ),

    label_full =
      paste0(
        short_name,
        "\n",
        n_cohorts
      )
  )
```

```{r africa-map, fig.width=14, fig.height=8}
map_max <-
  max(
    africa_facets$n_cohorts,
    na.rm = TRUE
  )


p_facets_map <-
  ggplot() +

  geom_sf(

    data =
      africa_facets,

    aes(
      fill =
        fill_value
    ),

    color =
      "grey45",

    linewidth =
      0.25
  ) +

  scale_fill_gradientn(

    colours =
      c(
        "#EAF3FB",
        "#9CCBEA",
        "#3B8BC2",
        "#004B7A"
      ),

    na.value =
      "grey92",

    limits =
      c(
        1,
        map_max
      ),

    name =
      "Number of cohorts"
  ) +

  ggrepel::geom_text_repel(

    data =
      label_df,

    aes(

      x =
        X,

      y =
        Y,

      label =
        label_full
    ),

    size =
      2.4,

    lineheight =
      0.88,

    box.padding =
      0.25,

    point.padding =
      0.10,

    min.segment.length =
      0,

    segment.color =
      "grey45",

    segment.size =
      0.20,

    max.overlaps =
      Inf,

    seed =
      123
  ) +

  coord_sf(

    xlim =
      c(
        -30,
        58
      ),

    ylim =
      c(
        -37,
        39
      ),

    expand =
      FALSE,

    clip =
      "off"
  ) +

  facet_wrap(

    ~panel,

    ncol =
      2,

    drop =
      FALSE
  ) +

  ggspatial::annotation_north_arrow(

    location =
      "tr",

    which_north =
      "true",

    height =
      grid::unit(
        0.75,
        "cm"
      ),

    width =
      grid::unit(
        0.75,
        "cm"
      ),

    style =
      ggspatial::north_arrow_fancy_orienteering
  ) +

  ggspatial::annotation_scale(

    location =
      "bl",

    width_hint =
      0.28,

    text_cex =
      0.60
  ) +

  theme_void(
    base_size =
      12
  ) +

  theme(

    strip.text =
      element_text(
        face = "bold",
        size = 11
      ),

    legend.position =
      "right",

    plot.margin =
      margin(
        5,
        15,
        5,
        15
      )
  )


p_facets_map
```

# 21. Country and region summaries

```{r country-region-summaries}
country_summary_total <-
  country_long %>%

  distinct(
    cohort_id,
    country,
    iso3
  ) %>%

  count(
    country,
    name =
      "Total cohorts"
  ) %>%

  arrange(
    desc(
      `Total cohorts`
    ),
    country
  )


country_summary_interface <-
  country_long %>%

  filter(
    !is.na(
      broad_interface_cleaned
    )
  ) %>%

  distinct(
    cohort_id,
    country,
    broad_interface_cleaned
  ) %>%

  count(
    country,
    broad_interface_cleaned,
    name =
      "n"
  ) %>%

  pivot_wider(

    names_from =
      broad_interface_cleaned,

    values_from =
      n,

    values_fill =
      0
  )


country_summary_table <-
  country_summary_total %>%

  left_join(
    country_summary_interface,
    by =
      "country"
  ) %>%

  rename(
    Country =
      country
  )


country_summary_table


country_long_region <-
  country_long %>%

  mutate(

    region =
      case_when(

        iso3 %in%
          c(
            "KEN",
            "UGA",
            "TZA",
            "ETH",
            "SDN",
            "SSD",
            "RWA",
            "BDI",
            "ERI",
            "SOM",
            "DJI",
            "SYC"
          ) ~
          "East Africa",

        iso3 %in%
          c(
            "GHA",
            "NGA",
            "GMB",
            "SEN",
            "MLI",
            "BFA",
            "NER",
            "BEN",
            "CIV",
            "GNB",
            "GIN",
            "LBR",
            "SLE",
            "TGO"
          ) ~
          "West Africa",

        iso3 %in%
          c(
            "ZAF",
            "ZWE",
            "ZMB",
            "MOZ",
            "MWI",
            "NAM",
            "BWA",
            "AGO",
            "MDG",
            "MUS",
            "LSO",
            "SWZ"
          ) ~
          "Southern Africa",

        iso3 %in%
          c(
            "COD",
            "COG",
            "CAF",
            "CMR",
            "TCD",
            "GAB",
            "GNQ"
          ) ~
          "Central Africa",

        iso3 %in%
          c(
            "EGY",
            "DZA",
            "MAR",
            "TUN",
            "LBY"
          ) ~
          "North Africa",

        TRUE ~
          "Other/unclear"
      )
  )


region_summary_table <-
  country_long_region %>%

  distinct(
    cohort_id,
    country,
    region
  ) %>%

  count(
    region,
    name =
      "Total cohort-country observations"
  ) %>%

  arrange(
    desc(
      `Total cohort-country observations`
    )
  )


region_summary_table
```

# 22. Recruitment strategies

```{r recruitment-strategies}
cohort_master <-
  cohort_master %>%

  mutate(

    recruitment_method_clean =
      recruitment_method %>%

      as.character() %>%

      str_to_lower() %>%

      str_replace_all(
        "–|—",
        "-"
      ) %>%

      str_squish(),

    recruitment_strategy_group =
      case_when(

        str_detect(
          recruitment_method_clean,
          "convenience"
        ) ~
          "Convenience sampling",

        str_detect(
          recruitment_method_clean,
          paste0(
            "facility|hospital|clinic|antenatal|prenatal|maternity|delivery|",
            "health centre|health-facility|clinical|routine clinical"
          )
        ) ~
          "Facility-based recruitment",

        str_detect(
          recruitment_method_clean,
          paste0(
            "pregnant women|newborn|children|birth cohort|school-based|",
            "male children|dog-bite|occupational|poultry|seasonal recruitment|",
            "administrative enrolment|retrospective extraction|prospective cohort"
          )
        ) ~
          "Specific population recruitment",

        str_detect(
          recruitment_method_clean,
          paste0(
            "random|cluster|stratified|multistage|multi-stage|systematic|",
            "representative|subsample|panel survey|sampling|spatial transects"
          )
        ) ~
          "Probability-based sampling",

        str_detect(
          recruitment_method_clean,
          paste0(
            "community|village|community health volunteers|chvs|parents|",
            "guardians|local|house-to-house"
          )
        ) ~
          "Community-based recruitment",

        str_detect(
          recruitment_method_clean,
          paste0(
            "household|census|enumeration|enumaration|population-based|",
            "population based|population census|population enumeration|hdss|",
            "surveillance|sample registration system|dynamic enrollment|",
            "longitudinal demographic|continuous demographic|open population|",
            "dynamic cohort|whole population|whole-population|",
            "complete population|entire village|door-to-door|door-door"
          )
        ) ~
          "Household / population census and enumeration",

        TRUE ~
          NA_character_
      )
  )


unclassified_recruitment <-
  cohort_master %>%

  filter(
    is.na(
      recruitment_strategy_group
    )
  ) %>%

  count(
    recruitment_method,
    sort =
      TRUE
  )


unclassified_recruitment


recruitment_summary <-
  cohort_master %>%

  filter(
    !is.na(
      recruitment_strategy_group
    )
  ) %>%

  count(
    recruitment_strategy_group,
    sort =
      TRUE
  ) %>%

  mutate(

    percent =
      round(
        100 *
          n /
          nrow(
            cohort_master
          ),
        1
      ),

    `N (%)` =
      paste0(
        n,
        " (",
        percent,
        "%)"
      )
  )


recruitment_summary_ft <-
  recruitment_summary %>%

  select(

    `Recruitment strategy` =
      recruitment_strategy_group,

    `N (%)`
  ) %>%

  flextable() %>%

  theme_booktabs() %>%

  autofit() %>%

  align(
    part =
      "all",
    align =
      "left"
  ) %>%

  align(
    j =
      "N (%)",
    align =
      "center",
    part =
      "all"
  ) %>%

  bold(
    part =
      "header"
  ) %>%

  set_caption(
    "Recruitment strategies of included cohorts"
  )


recruitment_summary_ft
```

# 23. Save final analysis datasets and key figures

```{r save-outputs}
write.csv(
  cohort_clean,
  "cohort_analysis_dataset.csv",
  row.names = FALSE
)


write.csv(
  infectious_outcome_counts,
  "infectious_outcome_summary.csv",
  row.names = FALSE
)


write.csv(
  model_results_clean,
  "data_integration_regression_results.csv",
  row.names = FALSE
)


ggsave(
  "figure_infectious_outcomes.png",
  p_infectious_outcomes,
  width = 12,
  height = 8,
  dpi = 300
)


ggsave(
  "figure_data_integration_forest.png",
  p_forest,
  width = 10,
  height = 7,
  dpi = 300
)


ggsave(
  "figure_africa_cohort_maps.png",
  p_facets_map,
  width = 14,
  height = 8,
  dpi = 300
)
```

# 24. Reproducibility information

```{r session-info}
sessionInfo()
```